---
title: "Multidimensional model all datasets"
author: Simone Colombara, Alessia Cotroneo, Francesco De Caro, Riccardo Morandi, Chiara   Schembri,
  Alfredo Zapiola
date: "2022-12-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 6,fig.height = 4)
library(tidyverse)
library(rayshader)
library(patchwork)
library(viridis)
library(plot.matrix)
library(bayesImageS)
library(stats)
library(ggplot2)
library(Rcpp)
library(RcppArmadillo)
library(coda)
library(MASS)
library(ellipse)
library(plyr)
library(ggExtra)
```

```{r, warning=FALSE}
sourceCpp("/Users/macbookpro/Documents/Bayesian Statistics/Project/Cpp_code/GibbsGMM.cpp")
sourceCpp("/Users/macbookpro/Documents/Bayesian Statistics/Project/Cpp_code/GibbsSampler_parallel.cpp")
sourceCpp("/Users/macbookpro/Documents/Bayesian Statistics/Project/Cpp_code/mcmcPotts.cpp")
```

## costructing the dataset with the first pca score of each dataset

Lipidi

```{r}
L = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-IM-Step1-Step2-Step4-Step5-101.txt")
L0 = L
L0[is.na(L0)] = 0
Lpixels = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-XYCoordinates-Step1-Step2-Step4-Step5-101.txt")
Lmax_n_of_pixel = c(157,178)
```

Glicani

```{r}
G = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/Glicani/85 variabili/101_glicani-PreProcessed-IM-Step1-Step2-Step4-Step5-101.txt")
G0 = G
G0[is.na(G0)] = 0
Gpixels = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/Glicani/85 variabili/101_glicani-PreProcessed-XYCoordinates-Step1-Step2-Step4-Step5-101.txt")
Gmax_n_of_pixel = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/Glicani/85 variabili/101_glicani-PreProcessed-maxXY-Step1-Step2-Step4-Step5-101.txt")
```

Peptidi

```{r}
P = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/Peptidi/154 variabili/101_peptidi-PreProcessed-IM-Step1-Step2-Step4-Step5-101.txt")
P0 = P
P0[is.na(P0)] = 0
Ppixels = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/Peptidi/154 variabili/101_peptidi-PreProcessed-XYCoordinates-Step1-Step2-Step4-Step5-101.txt")
Pmax_n_of_pixel = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/Peptidi/154 variabili/101_peptidi-PreProcessed-maxXY-Step1-Step2-Step4-Step5-101.txt")
```

calculate the first rpincipal components

```{r}
pcaL = princomp(L0)$scores[,1]
pcaG = princomp(G0)$scores[,1]
pcaP = princomp(P0)$scores[,1]
```

take care of the missing values

number of observations per pixels
obs: the max umber of pixels is the same for all three datasets
```{r}
obs  = matrix(0,Lmax_n_of_pixel[1],Lmax_n_of_pixel[2])
lipidi = matrix(NA,Lmax_n_of_pixel[1],Lmax_n_of_pixel[2])
peptidi = matrix(NA,Lmax_n_of_pixel[1],Lmax_n_of_pixel[2])
glicani = matrix(NA,Lmax_n_of_pixel[1],Lmax_n_of_pixel[2])


for(i in 1:dim(Lpixels)[1]){
  obs[Lpixels[i,1],Lpixels[i,2]] = obs[Lpixels[i,1],Lpixels[i,2]] + 1
  lipidi[Lpixels[i,1],Lpixels[i,2]] = pcaL[i]
}

for(i in 1:dim(Gpixels)[1]){
  obs[Gpixels[i,1],Gpixels[i,2]] = obs[Gpixels[i,1],Gpixels[i,2]] + 1
  glicani[Gpixels[i,1],Gpixels[i,2]] = pcaG[i]
}

for(i in 1:dim(Ppixels)[1]){
  obs[Ppixels[i,1],Ppixels[i,2]] = obs[Ppixels[i,1],Ppixels[i,2]] + 1
  peptidi[Ppixels[i,1],Ppixels[i,2]] = pcaP[i]
}

```

let us create the mask of the full observations

```{r}
mask = matrix(0,Lmax_n_of_pixel[1],Lmax_n_of_pixel[2])
for(i in 1:Lmax_n_of_pixel[1]){
  for( j in 1:Lmax_n_of_pixel[2]){
    if(obs[i,j]==3)
      mask[i,j] = 1
  }
}
nobs = table(mask)[2]
```


```{r}
data = matrix(0,3,nobs)
cnt = 1
x = vector(length = nobs)
y = vector(length =  nobs)
for(i in 1:Lmax_n_of_pixel[1]){
  for( j in 1:Lmax_n_of_pixel[2]){
    if(mask[i,j] == 1){
     data[1,cnt] = lipidi[i,j] 
     data[2,cnt] = glicani[i,j] 
     data[3,cnt] = peptidi[i,j] 
     x[cnt] = i
     y[cnt] = j
     cnt = cnt + 1
    }
  }
}
```

dataframe for plot and storing results

```{r}
df = data.frame("lipidi" = data[1,], "glicani" = data[2,], "peptidi" = data[3,], "x" = x, "y" = y)
```

# K = 3

## GMM

```{r}
k = 3
d = 3
B0 = diag(100,d,d)
V0 = diag(10,d,d)
n0 = 10

priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
priors$lambda <- rep(1,k)
```

```{r}
iter = 10000
burnin = 5000
```

```{r}
resGMM3 <- GibbsGMM(data,iter,burnin,priors)
```


```{r}
allocation = rep(0,dim(pixels)[1])
for (i in 1:dim(pixels)[1]){
  allocation[i] = which.max(resGMM3$alloc[i,])
}
df$allocGMM3 = allocation
```

```{r}
ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = as.factor(allocGMM3))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")
ggsave("allGMM3clust.png",width = 20,units = "cm")
```

```{r}
G12 = ggplot(data = df,aes(x = lipidi,y = glicani,color = as.factor(allocGMM3))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) +   labs(color = "cluster") + theme(legend.position = "none")

G13 = ggplot(data = df,aes(x = lipidi,y = peptidi,color = as.factor(allocGMM3))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1)+   labs(color = "cluster") + theme(legend.position = "none")

G23 = ggplot(data = df,aes(x = glicani,y = peptidi,color = as.factor(allocGMM3))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1)+  labs(color = "cluster") + theme(legend.position = "none")

G12 + G13 + G23

ggsave("allGMM3sscattera.png",width = 20,units = "cm")
```

```{r}
G1 = ggplot(data = df,aes(x = lipidi,color = as.factor(allocGMM3), fill = as.factor(allocGMM3),after_stat(density))) + geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G2 = ggplot(data = df,aes(x = glicani,color = as.factor(allocGMM3), fill = as.factor(allocGMM3),after_stat(density))) +  geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G3 = ggplot(data = df,aes(x = peptidi,color = as.factor(allocGMM3), fill = as.factor(allocGMM3),after_stat(density))) +  geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G1 + G2 + G3 
ggsave("allGMM3marginals.png",width = 20,units = "cm")
```

```{r}
muchain = data.frame("mu" = c(resGMM3$mu[1,1,], resGMM3$mu[2,1,],resGMM3$mu[3,1,], resGMM3$mu[1,2,], resGMM3$mu[2,2,],
                              resGMM3$mu[3,2,], resGMM3$mu[1,3,], resGMM3$mu[2,3,],resGMM3$mu[3,3,]),
                     "index" = rep(seq(1,iter),d*k),"cluster" = c(rep(1,d*iter),rep(2,d*iter),rep(3,d*iter)), 
                     "comp" = rep(c(rep(1,iter),rep(2,iter),rep(3,iter)),k))

ggplot(data = muchain,aes(x = index,y = mu,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") + facet_wrap(~comp,scales = "free") + theme(legend.position = "bottom") + labs(color = "cluster")
```

```{r,echo=FALSE}

sigma1chain = array(dim = c(d,d,iter))
sigma2chain = array(dim = c(d,d,iter))
sigma3chain = array(dim = c(d,d,iter))
for(i in 1:iter){
  sigma1chain[,,i] = resGMM3$sigma[i][[1]][,,1]
  sigma2chain[,,i] = resGMM3$sigma[i][[1]][,,2]
  sigma3chain[,,i] = resGMM3$sigma[i][[1]][,,3]
}

sigma1 = apply(sigma1chain, c(1,2), mean)
sigma2 = apply(sigma2chain, c(1,2), mean)
sigma3 = apply(sigma3chain, c(1,2), mean)

sigmachain = data.frame("sigma" = c(sigma1chain[1,1,], sigma1chain[1,2,],sigma1chain[1,3,], 
                                    sigma1chain[2,1,], sigma1chain[2,2,],sigma1chain[2,3,], 
                                    sigma1chain[3,1,], sigma1chain[3,2,],sigma1chain[3,3,],
                                    sigma2chain[1,1,], sigma2chain[1,2,],sigma2chain[1,3,], 
                                    sigma2chain[2,1,], sigma2chain[2,2,],sigma2chain[2,3,], 
                                    sigma2chain[3,1,], sigma2chain[3,2,],sigma2chain[3,3,], 
                                    sigma3chain[1,1,], sigma3chain[1,2,],sigma3chain[1,3,], 
                                    sigma3chain[2,1,], sigma3chain[2,2,],sigma3chain[2,3,], 
                                    sigma3chain[3,1,], sigma3chain[3,2,],sigma3chain[3,3,]),
                        "index" = rep(seq(1,iter),d*d*k),
                     "cluster" = c(rep(1,d*d*iter),rep(2,d*d*iter),rep(3,d*d*iter)), "comp" =
                       rep(c(rep(11,iter),rep(12,iter),rep(13,iter),
                             rep(21,iter),rep(22,iter),rep(23,iter),
                             rep(31,iter),rep(32,iter),rep(33,iter)),k))

```

```{r}
ggplot(data = sigmachain,aes(x = index,y = sigma,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + facet_wrap(~comp,scales = "free") + labs(color = "cluster")
```

we still need to plot the covariance autocorr functions

```{r}
wchain <- data.frame("w" = c(resGMM3$w[,1],resGMM3$w[,2],resGMM3$w[,3]), "index" = rep(seq(1,iter),k), 
                     "cluster" =c(rep(1,iter),rep(2,iter),rep(3,iter)))

ggplot(data = wchain,aes(x = index,y = w,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H")  +
  theme(legend.position = "bottom") + labs(color = "cluster")
```


```{r}
G12 = ggplot(data = df,aes(x = lipidi,y = glicani,color = as.factor(allocGMM3))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(allocGMM3),scales = "free") +   labs(color = "cluster") + theme(legend.position = "none")

G13 = ggplot(data = df,aes(x = lipidi,y = peptidi,color = as.factor(allocGMM3))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(allocGMM3),scales = "free") +   labs(color = "cluster") + theme(legend.position = "none")

G23 = ggplot(data = df,aes(x = glicani,y = peptidi,color = as.factor(allocGMM3))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(allocGMM3),scales = "free") +  labs(color = "cluster") + theme(legend.position = "none")

G12 + G13 + G23

ggsave("allGMM3sscatterb.png",width = 20,units = "cm")
```
let us build the potts model now

## Potts Model

we need to know the geometry, we do it through the mask 

```{r}
neigh <- getNeighbors(mask = mask, c(2,2,0,0))
block <- getBlocks(mask = mask, 2)
```

### Pseudolikelihood

```{r}
betacritic = log(1 + sqrt(k))
initmu = matrix(0,d,k)
initsigma = array(n0*V0,dim = c(d,d,k))

priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
priors$beta <- c(0,2*betacritic)

mh <- list(bandwidth=0.1)
```

```{r}
resultsMCMC3 <- MCMCPotts(data,neigh,block,iter,burnin,priors,mh)
```

```{r}
allocationmcmc = rep(0,nobs)
for (i in 1:nobs){
  allocationmcmc[i] = which.max(resultsMCMC3$alloc[i,])
  # just to match the labels to above
  if(allocationmcmc[i]== 3){
    allocationmcmc[i] = 1
  }
  else if(allocationmcmc[i]== 1){
    allocationmcmc[i] = 3
  }
}

df$allocmcmc3 = allocationmcmc
```

```{r}
ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocmcmc3))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")
#ggsave("mcmc_k3cluster.png",width = 20,units = "cm")
```

```{r}
sumchain <- data.frame("sum" = resultsMCMC3$sum[,1], "index" = seq(1,iter))
betachain <- data.frame("beta" = resultsMCMC3$beta[,1], "index" = seq(1,iter))

G1 = ggplot(data = sumchain,aes(x = index,y = sum)) + geom_line() 
G2 = ggplot(data = betachain,aes(x = index,y = beta)) + geom_line() + geom_hline(yintercept = betacritic,color = "red",)
G1 + G2
```

the results are not good, since the beta goes too high and makes the results very uniform

### Fixing beta to betacritic

```{r}
priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
```

```{r}
resultsGibbs3 <- GibbsPotts(data,betacritic,initmu,initsigma,neigh,block,priors,iter,burnin)
```

```{r}
allocationgibbs = rep(0,nobs)
for (i in 1:nobs){
  allocationgibbs[i] = which.max(resultsGibbs3$alloc[i,])
  # just to match the labels to above
  if(allocationgibbs[i]== 3){
    allocationgibbs[i] = 2
  }
  else if(allocationgibbs[i]== 2){
    allocationgibbs[i] = 3
  }
  
}
df$allocgibbs3= allocationgibbs
```

```{r}
ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocgibbs3))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")
ggsave("allgibbs3.png")
```


```{r}
G12 = ggplot(data = df,aes(x = lipidi,y = glicani,color = as.factor(allocgibbs3))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) +   labs(color = "cluster") + theme(legend.position = "none")

G13 = ggplot(data = df,aes(x = lipidi,y = peptidi,color = as.factor(allocgibbs3))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1)+   labs(color = "cluster") + theme(legend.position = "none")

G23 = ggplot(data = df,aes(x = glicani,y = peptidi,color = as.factor(allocgibbs3))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1)+  labs(color = "cluster") + theme(legend.position = "none")

G12 + G13 + G23

ggsave("allGibbs3scattera.png",width = 20,units = "cm")
```

```{r}
G1 = ggplot(data = df,aes(x = lipidi,color = as.factor(allocgibbs3), fill = as.factor(allocgibbs3),after_stat(density))) + geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G2 = ggplot(data = df,aes(x = glicani,color = as.factor(allocgibbs3), fill = as.factor(allocgibbs3),after_stat(density))) +  geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G3 = ggplot(data = df,aes(x = peptidi,color = as.factor(allocgibbs3), fill = as.factor(allocgibbs3),after_stat(density))) +  geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G1 + G2 + G3 
ggsave("allGibbs3marginals.png",width = 20,units = "cm")
```

```{r}
muchain = data.frame("mu" = c(resultsGibbs3$mu[1,1,], resultsGibbs3$mu[2,1,],resultsGibbs3$mu[3,1,], resultsGibbs3$mu[1,2,],
                              resultsGibbs3$mu[2,2,],resultsGibbs3$mu[3,2,], resultsGibbs3$mu[1,3,],
                              resultsGibbs3$mu[2,3,],resultsGibbs3$mu[3,3,]),
                     "index" = rep(seq(1,iter),d*k),"cluster" = c(rep(1,d*iter),rep(2,d*iter),rep(3,d*iter)), 
                     "comp" = rep(c(rep(1,iter),rep(2,iter),rep(3,iter)),k))

ggplot(data = muchain,aes(x = index,y = mu,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") + facet_wrap(~comp,scales = "free") + theme(legend.position = "bottom") + labs(color = "cluster")
```

```{r,echo=FALSE}

sigma1chain = array(dim = c(d,d,iter))
sigma2chain = array(dim = c(d,d,iter))
sigma3chain = array(dim = c(d,d,iter))
for(i in 1:iter){
  sigma1chain[,,i] = resultsGibbs3$sigma[i][[1]][,,1]
  sigma2chain[,,i] = resultsGibbs3$sigma[i][[1]][,,2]
  sigma3chain[,,i] = resultsGibbs3$sigma[i][[1]][,,3]
}

sigma1 = apply(sigma1chain, c(1,2), mean)
sigma2 = apply(sigma2chain, c(1,2), mean)
sigma3 = apply(sigma3chain, c(1,2), mean)

sigmachain = data.frame("sigma" = c(sigma1chain[1,1,], sigma1chain[1,2,],sigma1chain[1,3,], 
                                    sigma1chain[2,1,], sigma1chain[2,2,],sigma1chain[2,3,], 
                                    sigma1chain[3,1,], sigma1chain[3,2,],sigma1chain[3,3,],
                                    sigma2chain[1,1,], sigma2chain[1,2,],sigma2chain[1,3,], 
                                    sigma2chain[2,1,], sigma2chain[2,2,],sigma2chain[2,3,], 
                                    sigma2chain[3,1,], sigma2chain[3,2,],sigma2chain[3,3,], 
                                    sigma3chain[1,1,], sigma3chain[1,2,],sigma3chain[1,3,], 
                                    sigma3chain[2,1,], sigma3chain[2,2,],sigma3chain[2,3,], 
                                    sigma3chain[3,1,], sigma3chain[3,2,],sigma3chain[3,3,]),
                        "index" = rep(seq(1,iter),d*d*k),
                     "cluster" = c(rep(1,d*d*iter),rep(2,d*d*iter),rep(3,d*d*iter)), "comp" =
                       rep(c(rep(11,iter),rep(12,iter),rep(13,iter),
                             rep(21,iter),rep(22,iter),rep(23,iter),
                             rep(31,iter),rep(32,iter),rep(33,iter)),k))

```

```{r}
ggplot(data = sigmachain,aes(x = index,y = sigma,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + facet_wrap(~comp,scales = "free") + labs(color = "cluster")
```

we still need to plot the covariance autocorr functions

```{r}
G12 = ggplot(data = df,aes(x = lipidi,y = glicani,color = as.factor(allocgibbs3))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(allocgibbs3),scales = "free") +   labs(color = "cluster") + theme(legend.position = "none")

G13 = ggplot(data = df,aes(x = lipidi,y = peptidi,color = as.factor(allocgibbs3))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(allocgibbs3),scales = "free") +   labs(color = "cluster") + theme(legend.position = "none")

G23 = ggplot(data = df,aes(x = glicani,y = peptidi,color = as.factor(allocgibbs3))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(allocgibbs3),scales = "free") +  labs(color = "cluster") + theme(legend.position = "none")

G12 + G13 + G23

ggsave("allGibbs3sscatterb.png",width = 20,units = "cm")
```

let us look at the differences between the potts model and the GMM

```{r}
diff_gmm_gibbs = rep(0,nobs)
for (i in 1:nobs){
  if(df$allocGMM3[i]!=df$allocgibbs3[i]){
    diff_gmm_gibbs[i] = 1
  }
}
df$diff_gmm_gibbs = diff_gmm_gibbs
```

```{r}
G1 = ggplot(data = df,aes(x = lipidi,y = glicani,color = as.factor(diff_gmm_gibbs))) + geom_point(alpha = 0.5) +  scale_colour_viridis_d(option = "plasma") + theme(legend.position = "bottom") + labs(color = "differently classified")
G2 = ggplot(data = df,aes(x = lipidi,y = peptidi,color = as.factor(diff_gmm_gibbs))) + geom_point(alpha = 0.5) +  scale_colour_viridis_d(option = "plasma") + theme(legend.position = "bottom") + labs(color = "differently classified")
G3 = ggplot(data = df,aes(x = glicani,y = peptidi,color = as.factor(diff_gmm_gibbs))) + geom_point(alpha = 0.5) +  scale_colour_viridis_d(option = "plasma") + theme(legend.position = "bottom") + labs(color = "differently classified")

G1 + G2 +G3
ggsave("diff_gmm_gibbs_k3b.png")
```

```{r}
G1 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(diff_gmm_gibbs))) + scale_fill_viridis_d(option = "plasma") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "difference GMM Potts")
G2 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocGMM3))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G3 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocgibbs3))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G2 + G1 + G3
ggsave("diff_gmm_gibbs_k3.png")
```

# K = 6

## GMM

```{r}
k = 6
d = 3
B0 = diag(100,d,d)
V0 = diag(10,d,d)
n0 = 10

priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
priors$lambda <- rep(1,k)
```

```{r}
resGMM6 <- GibbsGMM(data,iter,burnin,priors)
```

```{r}
allocation = rep(0,nobs)
for (i in 1:nobs){
  allocation[i] = which.max(resGMM6$alloc[i,])
}
df$allocGMM6 = allocation
```

```{r}
ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = as.factor(allocGMM6))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")
ggsave("allGMM6clust.png",width = 20,units = "cm")
```

```{r}
G12 = ggplot(data = df,aes(x = lipidi,y = glicani,color = as.factor(allocGMM6))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) +   labs(color = "cluster") + theme(legend.position = "none")

G13 = ggplot(data = df,aes(x = lipidi,y = peptidi,color = as.factor(allocGMM6))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1)+   labs(color = "cluster") + theme(legend.position = "none")

G23 = ggplot(data = df,aes(x = glicani,y = peptidi,color = as.factor(allocGMM6))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1)+  labs(color = "cluster") + theme(legend.position = "none")

G12 + G13 + G23

ggsave("allGMM6scattera.png",width = 20,units = "cm")
```

```{r}
G1 = ggplot(data = df,aes(x = lipidi,color = as.factor(allocGMM6), fill = as.factor(allocGMM6),after_stat(density))) + geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G2 = ggplot(data = df,aes(x = glicani,color = as.factor(allocGMM6), fill = as.factor(allocGMM6),after_stat(density))) +  geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G3 = ggplot(data = df,aes(x = peptidi,color = as.factor(allocGMM6), fill = as.factor(allocGMM6),after_stat(density))) +  geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G1 + G2 + G3 
ggsave("allGMM6marginals.png",width = 20,units = "cm")
```


```{r}
muchain = data.frame("mu" = c(resGMM6$mu[1,1,], resGMM6$mu[2,1,],resGMM6$mu[3,1,],resGMM6$mu[1,2,], resGMM6$mu[2,2,], resGMM6$mu[3,2,],
                     resGMM6$mu[1,3,], resGMM6$mu[2,3,], resGMM6$mu[3,3,],resGMM6$mu[1,4,], resGMM6$mu[2,4,], resGMM6$mu[3,4,],
                     resGMM6$mu[1,5,], resGMM6$mu[2,5,], resGMM6$mu[3,5,],resGMM6$mu[1,6,], resGMM6$mu[2,6,], resGMM6$mu[3,4,]),
                     "index" = rep(seq(1,iter),d*k),
                     "cluster" = c(rep(1,d*iter),rep(2,d*iter),rep(3,d*iter),rep(4,d*iter),rep(5,d*iter),rep(6,d*iter)), 
                     "comp" = rep(c(rep(1,iter),rep(2,iter),rep(3,iter))),k)

ggplot(data = muchain,aes(x = index,y = mu,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") + facet_wrap(~comp,scales = "free") + theme(legend.position = "bottom") + labs(color = "cluster")
ggsave("gmm_k6muchain.png")
```

and for the covariance matrices

```{r}

sigma1chain = array(dim = c(d,d,iter))
sigma2chain = array(dim = c(d,d,iter))
sigma3chain = array(dim = c(d,d,iter))
sigma4chain = array(dim = c(d,d,iter))
sigma5chain = array(dim = c(d,d,iter))
sigma6chain = array(dim = c(d,d,iter))

for(i in 1:iter){
  sigma1chain[,,i] = resGMM6$sigma[i][[1]][,,1]
  sigma2chain[,,i] = resGMM6$sigma[i][[1]][,,2]
  sigma3chain[,,i] = resGMM6$sigma[i][[1]][,,3]
  sigma4chain[,,i] = resGMM6$sigma[i][[1]][,,4]
  sigma5chain[,,i] = resGMM6$sigma[i][[1]][,,5]
  sigma6chain[,,i] = resGMM6$sigma[i][[1]][,,6]
}

sigma1 = apply(sigma1chain, c(1,2), mean)
sigma2 = apply(sigma2chain, c(1,2), mean)
sigma3 = apply(sigma3chain, c(1,2), mean)
sigma4 = apply(sigma4chain, c(1,2), mean)
sigma5 = apply(sigma5chain, c(1,2), mean)
sigma6 = apply(sigma5chain, c(1,2), mean)

sigmachain = data.frame("sigma" = c(sigma1chain[1,1,], sigma1chain[1,2,],sigma1chain[1,3,], 
                                    sigma1chain[2,1,], sigma1chain[2,2,],sigma1chain[2,3,], 
                                    sigma1chain[3,1,], sigma1chain[3,2,],sigma1chain[3,3,],
                                    sigma2chain[1,1,], sigma2chain[1,2,],sigma2chain[1,3,], 
                                    sigma2chain[2,1,], sigma2chain[2,2,],sigma2chain[2,3,], 
                                    sigma2chain[3,1,], sigma2chain[3,2,],sigma2chain[3,3,], 
                                    sigma3chain[1,1,], sigma3chain[1,2,],sigma3chain[1,3,], 
                                    sigma3chain[2,1,], sigma3chain[2,2,],sigma3chain[2,3,], 
                                    sigma3chain[3,1,], sigma3chain[3,2,],sigma3chain[3,3,],
                                    sigma4chain[1,1,], sigma4chain[1,2,],sigma4chain[1,3,], 
                                    sigma4chain[2,1,], sigma4chain[2,2,],sigma4chain[2,3,], 
                                    sigma4chain[3,1,], sigma4chain[3,2,],sigma4chain[3,3,],
                                    sigma5chain[1,1,], sigma5chain[1,2,],sigma5chain[1,3,], 
                                    sigma5chain[2,1,], sigma5chain[2,2,],sigma5chain[2,3,], 
                                    sigma5chain[3,1,], sigma5chain[3,2,],sigma5chain[3,3,], 
                                    sigma6chain[1,1,], sigma6chain[1,2,],sigma6chain[1,3,], 
                                    sigma6chain[2,1,], sigma6chain[2,2,],sigma6chain[2,3,], 
                                    sigma6chain[3,1,], sigma6chain[3,2,],sigma6chain[3,3,]),
                        "index" = rep(seq(1,iter),d*d*k),
                     "cluster" = c(rep(1,d*d*iter),rep(2,d*d*iter),rep(3,d*d*iter),rep(4,d*d*iter),rep(5,d*d*iter),rep(6,d*d*iter)), 
                     "comp" =rep(c(rep(11,iter),rep(12,iter),rep(13,iter),
                             rep(21,iter),rep(22,iter),rep(23,iter),
                             rep(31,iter),rep(32,iter),rep(33,iter)),k))
```

```{r}
ggplot(data = sigmachain,aes(x = index,y = sigma,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + facet_wrap(~comp,scales = "free") + labs(color = "cluster") 
ggsave("gmm_k6sigmachain.png")
```

wee still need to plot the covariance autocorr functions

```{r}
wchain <- data.frame("w" = c(resGMM6$w[,1],resGMM6$w[,2],resGMM6$w[,3],resGMM6$w[,4],resGMM6$w[,5],resGMM6$w[,6]), 
                     "index" = rep(seq(1,iter),k), 
                     "cluster" =c(rep(1,iter),rep(2,iter),rep(3,iter),rep(4,iter),rep(5,iter),rep(6,iter)))

ggplot(data = wchain,aes(x = index,y = w,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H")  +
  theme(legend.position = "bottom") + labs(color = "cluster")
ggsave("gmm_k6wchain.png")
```

## Potts Model

### Pseudolikelihood

```{r}
betacritic = log(1 + sqrt(k))
initmu = matrix(0,d,k)
initsigma = array(n0*V0,dim = c(d,d,k))

priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
priors$beta <- c(0,2*betacritic)

mh <- list(bandwidth=0.1)
```

```{r}
resultsMCMC6 <- MCMCPotts(data,neigh,block,iter,burnin,priors,mh)
```

```{r}
allocationmcmc = rep(0,nobs)
for (i in 1:nobs){
  allocationmcmc[i] = which.max(resultsMCMC6$alloc[i,])
  # just to match the labels to above
  if(allocationmcmc[i]== 3){
    allocationmcmc[i] = 1
  }
  else if(allocationmcmc[i]== 1){
    allocationmcmc[i] = 3
  }
}

df$allocmcmc6 = allocationmcmc
```

```{r}
ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocmcmc6))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")
#ggsave("mcmc_k3cluster.png",width = 20,units = "cm")
```

```{r}
sumchain <- data.frame("sum" = resultsMCMC6$sum[,1], "index" = seq(1,iter))
betachain <- data.frame("beta" = resultsMCMC6$beta[,1], "index" = seq(1,iter))

G1 = ggplot(data = sumchain,aes(x = index,y = sum)) + geom_line() 
G2 = ggplot(data = betachain,aes(x = index,y = beta)) + geom_line() + geom_hline(yintercept = betacritic,color = "red",)
G1 + G2
```

the results are not good, since the beta goes too high and makes the results very uniform

### Fixing beta to betacritic

```{r}
priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
```

```{r}
resultsGibbs6 <- GibbsPotts(data,betacritic,initmu,initsigma,neigh,block,priors,iter,burnin)
```

```{r}
allocationgibbs = rep(0,nobs)
for (i in 1:nobs){
  allocationgibbs[i] = which.max(resultsGibbs6$alloc[i,])
  # just to match the labels to above
  if(allocationgibbs[i]== 1){
    allocationgibbs[i] = 6
  }
  else if(allocationgibbs[i]== 5){
    allocationgibbs[i] = 1
  }
  else if(allocationgibbs[i]== 3){
    allocationgibbs[i] = 2
  }
  else if(allocationgibbs[i]== 6){
    allocationgibbs[i] =3
  }
  else if(allocationgibbs[i]== 2){
    allocationgibbs[i] = 5
  }
}
df$allocgibbs6= allocationgibbs
```

```{r}
ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocgibbs6))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")
ggsave("allgibbs6.png")
```


```{r}
G12 = ggplot(data = df,aes(x = lipidi,y = glicani,color = as.factor(allocgibbs6))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) +   labs(color = "cluster") + theme(legend.position = "none")

G13 = ggplot(data = df,aes(x = lipidi,y = peptidi,color = as.factor(allocgibbs6))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1)+   labs(color = "cluster") + theme(legend.position = "none")

G23 = ggplot(data = df,aes(x = glicani,y = peptidi,color = as.factor(allocgibbs6))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1)+  labs(color = "cluster") + theme(legend.position = "none")

G12 + G13 + G23

ggsave("allGibbs6scattera.png",width = 20,units = "cm")
```

```{r}
G1 = ggplot(data = df,aes(x = lipidi,color = as.factor(allocgibbs6), fill = as.factor(allocgibbs6),after_stat(density))) + geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G2 = ggplot(data = df,aes(x = glicani,color = as.factor(allocgibbs6), fill = as.factor(allocgibbs6),after_stat(density))) +  geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G3 = ggplot(data = df,aes(x = peptidi,color = as.factor(allocgibbs6), fill = as.factor(allocgibbs6),after_stat(density))) +  geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G1 + G2 + G3 
ggsave("allGibbs6marginals.png",width = 20,units = "cm")
```

```{r}
muchain = data.frame("mu" = c(resultsGibbs6$mu[1,1,], resultsGibbs6$mu[2,1,],resultsGibbs6$mu[3,1,],resultsGibbs6$mu[1,2,],
                              resultsGibbs6$mu[2,2,], resultsGibbs6$mu[3,2,],resultsGibbs6$mu[1,3,], resultsGibbs6$mu[2,3,],
                              resultsGibbs6$mu[3,3,],resultsGibbs6$mu[1,4,], resultsGibbs6$mu[2,4,], resultsGibbs6$mu[3,4,],
                              resultsGibbs6$mu[1,5,], resultsGibbs6$mu[2,5,], resultsGibbs6$mu[3,5,],resultsGibbs6$mu[1,6,],
                              resultsGibbs6$mu[2,6,], resultsGibbs6$mu[3,4,]),
                     "index" = rep(seq(1,iter),d*k),
                     "cluster" = c(rep(1,d*iter),rep(2,d*iter),rep(3,d*iter),rep(4,d*iter),rep(5,d*iter),rep(6,d*iter)), 
                     "comp" = rep(c(rep(1,iter),rep(2,iter),rep(3,iter))),k)

ggplot(data = muchain,aes(x = index,y = mu,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") + facet_wrap(~comp,scales = "free") + theme(legend.position = "bottom") + labs(color = "cluster")
ggsave("gibbs_k6muchain.png")
```

```{r,echo=FALSE}

sigma1chain = array(dim = c(d,d,iter))
sigma2chain = array(dim = c(d,d,iter))
sigma3chain = array(dim = c(d,d,iter))
for(i in 1:iter){
  sigma1chain[,,i] = resultsGibbs3$sigma[i][[1]][,,1]
  sigma2chain[,,i] = resultsGibbs3$sigma[i][[1]][,,2]
  sigma3chain[,,i] = resultsGibbs3$sigma[i][[1]][,,3]
}

sigma1 = apply(sigma1chain, c(1,2), mean)
sigma2 = apply(sigma2chain, c(1,2), mean)
sigma3 = apply(sigma3chain, c(1,2), mean)

sigmachain = data.frame("sigma" = c(sigma1chain[1,1,], sigma1chain[1,2,],sigma1chain[1,3,], 
                                    sigma1chain[2,1,], sigma1chain[2,2,],sigma1chain[2,3,], 
                                    sigma1chain[3,1,], sigma1chain[3,2,],sigma1chain[3,3,],
                                    sigma2chain[1,1,], sigma2chain[1,2,],sigma2chain[1,3,], 
                                    sigma2chain[2,1,], sigma2chain[2,2,],sigma2chain[2,3,], 
                                    sigma2chain[3,1,], sigma2chain[3,2,],sigma2chain[3,3,], 
                                    sigma3chain[1,1,], sigma3chain[1,2,],sigma3chain[1,3,], 
                                    sigma3chain[2,1,], sigma3chain[2,2,],sigma3chain[2,3,], 
                                    sigma3chain[3,1,], sigma3chain[3,2,],sigma3chain[3,3,],
                                    sigma4chain[1,1,], sigma4chain[1,2,],sigma4chain[1,3,], 
                                    sigma4chain[2,1,], sigma4chain[2,2,],sigma4chain[2,3,], 
                                    sigma4chain[3,1,], sigma4chain[3,2,],sigma4chain[3,3,],
                                    sigma5chain[1,1,], sigma5chain[1,2,],sigma5chain[1,3,], 
                                    sigma5chain[2,1,], sigma5chain[2,2,],sigma5chain[2,3,], 
                                    sigma5chain[3,1,], sigma5chain[3,2,],sigma5chain[3,3,], 
                                    sigma6chain[1,1,], sigma6chain[1,2,],sigma6chain[1,3,], 
                                    sigma6chain[2,1,], sigma6chain[2,2,],sigma6chain[2,3,], 
                                    sigma6chain[3,1,], sigma6chain[3,2,],sigma6chain[3,3,]),
                        "index" = rep(seq(1,iter),d*d*k),
                     "cluster" = c(rep(1,d*d*iter),rep(2,d*d*iter),rep(3,d*d*iter),rep(4,d*d*iter),rep(5,d*d*iter),rep(6,d*d*iter)), 
                     "comp" =rep(c(rep(11,iter),rep(12,iter),rep(13,iter),
                             rep(21,iter),rep(22,iter),rep(23,iter),
                             rep(31,iter),rep(32,iter),rep(33,iter)),k))

```

```{r}
ggplot(data = sigmachain,aes(x = index,y = sigma,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + facet_wrap(~comp,scales = "free") + labs(color = "cluster")
ggsave("gibbs_k6sigmachain.png")
```

we still need to plot the covariance autocorr functions

```{r}
G12 = ggplot(data = df,aes(x = lipidi,y = glicani,color = as.factor(allocgibbs6))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(allocgibbs6),scales = "free") +   labs(color = "cluster") + theme(legend.position = "none")

G13 = ggplot(data = df,aes(x = lipidi,y = peptidi,color = as.factor(allocgibbs6))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(allocgibbs6),scales = "free") +   labs(color = "cluster") + theme(legend.position = "none")

G23 = ggplot(data = df,aes(x = glicani,y = peptidi,color = as.factor(allocgibbs6))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") + theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(allocgibbs6),scales = "free") +  labs(color = "cluster") + theme(legend.position = "none")

G12 + G13 + G23

ggsave("allGibbs6sscatterb.png",width = 20,units = "cm")
```

let us look at the differences between the potts model and the GMM

```{r}
diff_gmm_gibbs = rep(0,nobs)
for (i in 1:nobs){
  if(df$allocGMM6[i]!=df$allocgibbs6[i]){
    diff_gmm_gibbs[i] = 1
  }
}
df$diff_gmm_gibbs6 = diff_gmm_gibbs
```

```{r}
G1 = ggplot(data = df,aes(x = lipidi,y = glicani,color = as.factor(diff_gmm_gibbs6))) + geom_point(alpha = 0.5) +  scale_colour_viridis_d(option = "plasma") + theme(legend.position = "bottom") + labs(color = "differently classified")
G2 = ggplot(data = df,aes(x = lipidi,y = peptidi,color = as.factor(diff_gmm_gibbs6))) + geom_point(alpha = 0.5) +  scale_colour_viridis_d(option = "plasma") + theme(legend.position = "bottom") + labs(color = "differently classified")
G3 = ggplot(data = df,aes(x = glicani,y = peptidi,color = as.factor(diff_gmm_gibbs6))) + geom_point(alpha = 0.5) +  scale_colour_viridis_d(option = "plasma") + theme(legend.position = "bottom") + labs(color = "differently classified")

G1 + G2 +G3
ggsave("diff_gmm_gibbs_k6b.png")
```

```{r}
G1 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(diff_gmm_gibbs6))) + scale_fill_viridis_d(option = "plasma") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "difference GMM Potts")
G2 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocGMM6))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G3 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocgibbs6))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G2 + G1 + G3
ggsave("diff_gmm_gibbs_k6.png")
```

the difference is not negligeble in this case
