---
title: "Sparse Finite Mixtures"
author: Simone Colombara, Alessia Cotroneo, Francesco De Caro, Riccardo Morandi, Chiara
  Schembri, Alfredo Zapiola
date: "2022-11-04"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plot.matrix)
library(bayesImageS)
library(stats)
library(viridis)
```

```{r}
D = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-IM-Step1-Step2-Step4-Step5-101.txt")
D0 = D
D0[is.na(D0)] = 0

pixels = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-XYCoordinates-Step1-Step2-Step4-Step5-101.txt")
colnames(D0) = substr(colnames(D0),1,4)
colnames(pixels) = c("x","y")
Data_long = as_tibble(data.frame( pixels, D0 ))
max_number_of_pixels = apply(Data_long[,1:2],2,max)
```

## CONVENTOIONAL PCA ON VECTOR DATA TO BE UPDATED ONCE WE FIX FPCA

```{r}
pca = princomp(D0)
```

auxiliary vector of principal components

```{r}
pcascore1vec = pca$scores[,1]
```

```{r}
par(mfrow=c(1,2))
hist(pcascore1vec,main = "first pc score",breaks = 100)
plot(density(pcascore1vec),main = "kernel density first pc score")
```



# all prior means are set to zero

## k = 6

the idea is to use a large k and small alpha in order to assign more weights to the edges of the symplex trying to induce sparsity

```{r}
q <- 5

priorsSGMM5 <- list()
priorsSGMM5$k <- q
priorsSGMM5$lambda <- rep(0.01,q)
priorsSGMM5$mu <- rep(0,q)
priorsSGMM5$mu.sd <- rep(100,q)
priorsSGMM5$sigma <- rep(2,q)
priorsSGMM5$sigma.nu <- rep(1.5,q)

resSGMM5 <- gibbsGMM(pcascore1vec, niter = 1000, nburn = 500, priors = priorsSGMM5)

clusteringSGMM5 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clusteringSGMM5[pixels[i,1],pixels[i,2]] = which.max(resSGMM5$alloc[i,])
}

print(table(clusteringSGMM5))
```

```{r}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clusteringSGMM5, border=NA,asp = TRUE,col =topo.colors(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```


## k = 10

the idea is to use a large k and small alpha in order to assign more weights to the edges of the symplex trying to induce sparsity 

```{r}
q <- 10

priorsSGMM10 <- list()
priorsSGMM10$k <- q
priorsSGMM10$lambda <- rep(0.01,q)
priorsSGMM10$mu <- rep(0,q)
priorsSGMM10$mu.sd <- rep(100,q)
priorsSGMM10$sigma <- rep(2,q)
priorsSGMM10$sigma.nu <- rep(1.5,q)

resSGMM10 <- gibbsGMM(pcascore1vec, niter = 1000, nburn = 500, priors = priorsSGMM10)

clusteringSGMM10 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clusteringSGMM10[pixels[i,1],pixels[i,2]] = which.max(resSGMM10$alloc[i,])
}

print(table(clusteringSGMM10))
```

```{r}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clusteringSGMM10, border=NA,asp = TRUE,col= topo.colors(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```



## k = 15

```{r}
q <- 15

priorsSGMM15 <- list()
priorsSGMM15$k <- q
priorsSGMM15$lambda <- rep(0.01,q)
priorsSGMM15$mu <- rep(0,q)
priorsSGMM15$mu.sd <- rep(100,q)
priorsSGMM15$sigma <- rep(2,q)
priorsSGMM15$sigma.nu <- rep(1.5,q)

resSGMM15 <- gibbsGMM(pcascore1vec, niter = 1000, nburn = 500, priors = priorsSGMM15)

clusteringSGMM15 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clusteringSGMM15[pixels[i,1],pixels[i,2]] = which.max(resSGMM15$alloc[i,])
}

print(table(clusteringSGMM15))
```

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clusteringSGMM15, border=NA,asp = TRUE,col =topo.colors(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```



## k = 20

```{r}
q <- 20

priorsSGMM20 <- list()
priorsSGMM20$k <- q
priorsSGMM20$lambda <- rep(0.001,q)
priorsSGMM20$mu <- rep(0,q)
priorsSGMM20$mu.sd <- rep(100,q)
priorsSGMM20$sigma <- rep(2,q)
priorsSGMM20$sigma.nu <- rep(1.5,q)

resSGMM20 <- gibbsGMM(pcascore1vec, niter = 1000, nburn = 500, priors = priorsSGMM20)

clusteringSGMM20 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clusteringSGMM20[pixels[i,1],pixels[i,2]] = which.max(resSGMM20$alloc[i,])
}

print(table(clusteringSGMM20))
```

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clusteringSGMM20, border=NA,asp = TRUE,col =topo.colors(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```

# Prior means spread out

## k = 5

the idea is to use a large k and small alpha in order to assign more weights to the edges of the symplex trying to induce sparsity

```{r}
q <- 5

priorsSGMM5 <- list()
priorsSGMM5$k <- q
priorsSGMM5$lambda <- rep(0.01,q)
priorsSGMM5$mu <- seq(from = -30, to = 20, length.out = q)
priorsSGMM5$mu.sd <- rep(100,q)
priorsSGMM5$sigma <- rep(2,q)
priorsSGMM5$sigma.nu <- rep(1.5,q)

resSGMM5 <- gibbsGMM(pcascore1vec, niter = 1000, nburn = 500, priors = priorsSGMM5)

clusteringSGMM5 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clusteringSGMM5[pixels[i,1],pixels[i,2]] = which.max(resSGMM5$alloc[i,])
}

print(table(clusteringSGMM5))
```

```{r}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clusteringSGMM5, border=NA,asp = TRUE,col =topo.colors(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```


## k = 10

the idea is to use a large k and small alpha in order to assign more weights to the edges of the symplex trying to induce sparsity 

```{r}
q <- 10

priorsSGMM10 <- list()
priorsSGMM10$k <- q
priorsSGMM10$lambda <- rep(0.01,q)
priorsSGMM10$mu <- seq(from = -30, to = 20, length.out = q)
priorsSGMM10$mu.sd <- rep(100,q)
priorsSGMM10$sigma <- rep(2,q)
priorsSGMM10$sigma.nu <- rep(1.5,q)

resSGMM10 <- gibbsGMM(pcascore1vec, niter = 1000, nburn = 500, priors = priorsSGMM10)

clusteringSGMM10 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clusteringSGMM10[pixels[i,1],pixels[i,2]] = which.max(resSGMM10$alloc[i,])
}

print(table(clusteringSGMM10))
```

```{r}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clusteringSGMM10, border=NA,asp = TRUE,col= topo.colors(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```



## k = 15

```{r}
q <- 15

priorsSGMM15 <- list()
priorsSGMM15$k <- q
priorsSGMM15$lambda <- rep(0.01,q)
priorsSGMM15$mu <- seq(from = -30, to = 20, length.out = q)
priorsSGMM15$mu.sd <- rep(100,q)
priorsSGMM15$sigma <- rep(2,q)
priorsSGMM15$sigma.nu <- rep(1.5,q)

resSGMM15 <- gibbsGMM(pcascore1vec, niter = 1000, nburn = 500, priors = priorsSGMM15)

clusteringSGMM15 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clusteringSGMM15[pixels[i,1],pixels[i,2]] = which.max(resSGMM15$alloc[i,])
}

print(table(clusteringSGMM15))
```

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clusteringSGMM15, border=NA,asp = TRUE,col =topo.colors(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```



## k = 20

```{r}
q <- 20

priorsSGMM20 <- list()
priorsSGMM20$k <- q
priorsSGMM20$lambda <- rep(0.001,q)
priorsSGMM20$mu <- seq(from = -34, to = 20, length.out = q)
priorsSGMM20$mu.sd <- rep(100,q)
priorsSGMM20$sigma <- rep(2,q)
priorsSGMM20$sigma.nu <- rep(1.5,q)

resSGMM20 <- gibbsGMM(pcascore1vec, niter = 1000, nburn = 500, priors = priorsSGMM20)

clusteringSGMM20 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clusteringSGMM20[pixels[i,1],pixels[i,2]] = which.max(resSGMM20$alloc[i,])
}

print(table(clusteringSGMM20))
```

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clusteringSGMM20, border=NA,asp = TRUE,col =topo.colors(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```
