---
title: "Data Inspection Glicani"
author: Simone Colombara, Alessia Cotroneo, Francesco De Caro, Riccardo Morandi, Chiara   Schembri,
  Alfredo Zapiola
date: "2022-11-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 8,fig.height = 6)
library(tidyverse)
library(rayshader)
library(patchwork)
library(skimr)
library(visdat)
library(ggplot2)
library(GGally)
library(corrplot)
```

# A first inspection of the dataset with 84 variables

## importing the data


```{r, read data}
G = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/Glicani/85 variabili/101_glicani-PreProcessed-IM-Step1-Step2-Step4-Step5-101.txt")
sum(is.na(G))
```
 the numbers of na is substantial
```{r }
 vis_miss(G,warn_large_data = FALSE)
```
the missing data is about 56% 
 
```{r}
skim(G)
```

we observe that the missing data is not uniform in the mz, there are some values for which only 20 - 30% of the pixel have a value, and this tends to be small

we replace the missing data with 0 since it means the data for that mz was under threshold


```{r}
G0 = G
G0[is.na(G0)] = 0
```

# correlation matrix
```{r}
cm <- cor(G0)
colnames(G0) = substr(colnames(G0),1,4)
corrplot(cm, method = "color", tl.pos = 'n')
```

the correlation is not high between the features, this is different from the Lipids

# preliminary plotts

```{r}
pixels = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/Glicani/85 variabili/101_glicani-PreProcessed-XYCoordinates-Step1-Step2-Step4-Step5-101.txt")
colnames(G0) = substr(colnames(G0),1,4)
colnames(pixels) = c("x","y")
max_n_of_pixel = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/Glicani/85 variabili/101_glicani-PreProcessed-maxXY-Step1-Step2-Step4-Step5-101.txt")
```

```{r}
Data_long            = as_tibble(data.frame( pixels, G0 ))
max_number_of_pixels = apply(Data_long[,1:2],2,max)

Data_array = matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])

Data_array = array(NA,c(max_number_of_pixels[1],max_number_of_pixels[2],ncol(G0)))

sum(is.na(G0))

# there must be a better way to do this
for(k in 1:ncol(G0)){
  for(i in 1:nrow(Data_long)){
  Data_array[Data_long$x[i],Data_long$y[i],k] = G0[i,k]
  }
}

dim(Data_array)

Data_very_long = reshape2::melt(Data_long,c("x","y")) %>% mutate(pixel_ind = paste0(x,"_",y), value_ind = rep(1:nrow(Data_long),ncol(G0)))
Data_very_long = Data_very_long %>% group_by(pixel_ind) %>% mutate(n = row_number()) %>% ungroup() %>% mutate(mz = as.numeric(substr(variable,2,4)))

```


```{r}
Data_very_long = reshape2::melt(Data_long,c("x","y")) %>% mutate(pixel_ind = paste0(x,"_",y), value_ind = rep(1:nrow(Data_long),ncol(G0)))

Data_very_long = Data_very_long %>% group_by(pixel_ind) %>% mutate(n = row_number()) %>% ungroup() %>% mutate(mz = as.numeric(substr(variable,2,4)))


# subsampling to get a faster plot and not drain memory
sub_ind = sample(unique(Data_very_long$pixel_ind),1000)
# just to get the gist:
ggplot(Data_very_long %>% filter(pixel_ind %in% sub_ind))+
  geom_path(aes(x = mz, y = value, 
                col=pixel_ind, 
                group = pixel_ind),alpha=.5)+theme_bw()+theme(legend.position = "none")+xlab("m.z")+scale_color_viridis_d(option = "A")+
  scale_x_continuous(n.breaks = 20)
```

# investigating the different peaks

## first interval 
the first values are quite noisy

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X100))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X102))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X105))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X108))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X109))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X109.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```



```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X109.2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X110))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X111))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X112))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X112.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X112.2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X113))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X113.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X115))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X115.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X115.2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X116))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```

## first peak

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X117))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X117.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X119))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X125))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X125.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X127))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```

here we start seeing the spikes, they are for 125 a few spots the rest is quite uniform, for 127 we have it quite spread out

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X127))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X127.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X127.2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X129))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X129.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X131.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```

there are some holes in the data, the only structure is the one in 127


## noisy intervall 

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X133))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X135))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X135.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X112))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X135.2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X139))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```

still quite noisy with some spots with high values

## seond peak 

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X139))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X139.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X141))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X143))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X143.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X146))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```

we can see some struture in 143 same struture as before

## third peak

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X150))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X151))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X160))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X160.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X164))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X172))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```

150 is another level where wee have features the rest is just noise

## noisy intervall

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X190))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X190.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X192))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X192.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X192.2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X195))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```

just noise and low values, the same holes repeats the next is still just noise 

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X221))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X221.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X221.2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X230))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X231))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X231.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X232))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X237))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X248))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X252))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X252.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X252.2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```

this is just noise up to the end of the available mz


# let us compare the peaks 

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X109))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X127))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X127.1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X127.2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X143))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X150))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```

it is the same peaks with some degraded points

# strange patterns

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X113))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X125))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X129))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")


P1+P2+P3

```

there are this two strange patterns, the conentrates spikes anf the holes

# PCA on vecotr data (NON Functional)

```{r}
pca = princomp(G0)
plot(pca)
summary(pca)
```
we nee much more components to get a good  percentage of the variance than in the lipid case

```{r}

PCA1 = ggplot(Data_long %>% mutate(pca1 = pca$scores[,1]))+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = pca1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
PCA2 = ggplot(Data_long %>% mutate(pca2 = pca$scores[,2]))+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = pca2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
PCA3 = ggplot(Data_long %>% mutate(pca1 = pca$scores[,3]))+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = pca1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
PCA4 = ggplot(Data_long %>% mutate(pca2 = pca$scores[4]))+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = pca2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
PCA5 = ggplot(Data_long %>% mutate(pca1 = pca$scores[,5]))+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = pca1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
PCA6 = ggplot(Data_long %>% mutate(pca2 = pca$scores[,6]))+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = pca2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")


PCA1+PCA2+PCA3+PCA4+PCA5+PCA6
```

# comparing thi with the peaks

```{r}
PCA1 = ggplot(Data_long %>% mutate(pca1 = pca$scores[,1]))+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = pca1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
PCA2 = ggplot(Data_long %>% mutate(pca2 = pca$scores[,2]))+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = pca2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
PCA3 = ggplot(Data_long %>% mutate(pca1 = pca$scores[,3]))+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = pca1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P1 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X143))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X127))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X127.2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")


PCA1+PCA2+PCA3+P1+P2+P3
```

# Plotting the principal components

## first component

```{r,echo = FALSE}
mz_value = c(100,102,105,108,109,109.1,109.2,110,111,112,112.1,112.2,113,113.1,115,115.1,115.2,116,117,117.1,119,125,125.1,127,127.1,127.2,129,129.1,131,131.1,133,135,135.1,135.2,139,139.1,141,143,143.1,146,150,151,160,160.1,164,172,190,190.1,192,192.1,192.2,195,195.1,196,199,199.2,199.2,201,202,206,211,211.1,211.2,213,213.1,213.2,215,217,217.1,221,221.1,221.2,230,231,231.1,232,237,248,252,252.1,252.2,262,262.1,262.2)
b1 = data.frame(pca$loadings[,1])
b1$mz =mz_value
ggplot(b1)+ geom_path(aes(x = mz, y = pca.loadings...1.))+theme_bw()+theme(legend.position = "none")+xlab("m.z")+scale_color_viridis_d(option = "A")+
  scale_x_continuous(n.breaks = 20)
```
this are the three clear peaks that one can see in the data

## second component

```{r,echo = FALSE}
b2 = data.frame(pca$loadings[,2])
b2$mz = mz_value
ggplot(b2)+ geom_path(aes(x = mz, y = pca.loadings...2.))+theme_bw()+theme(legend.position = "none")+xlab("m.z")+scale_color_viridis_d(option = "A")+
  scale_x_continuous(n.breaks = 20)

```

## third component

```{r,echo = FALSE}
b2 = data.frame(pca$loadings[,3])
b2$mz = mz_value
ggplot(b2)+ geom_path(aes(x = mz, y = pca.loadings...3.))+theme_bw()+theme(legend.position = "none")+xlab("m.z")+scale_color_viridis_d(option = "A")+
  scale_x_continuous(n.breaks = 20)

```
