---
title: "Lipidi Multidimensional models investigation"
author: Simone Colombara, Alessia Cotroneo, Francesco De Caro, Riccardo Morandi, Chiara   Schembri,
  Alfredo Zapiola
date: "2022-12-22"
output:   
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 7,fig.height = 5)
library(tidyverse)
library(rayshader)
library(patchwork)
library(viridis)
library(plot.matrix)
library(bayesImageS)
library(stats)
library(ggplot2)
library(Rcpp)
library(RcppArmadillo)
library(coda)
library(MASS)
library(ellipse)
library(plyr)
library(ggExtra)
```

```{r, warning=FALSE}
sourceCpp("/Users/macbookpro/Documents/Bayesian Statistics/Project/Cpp_code/GibbsGMM.cpp")
sourceCpp("/Users/macbookpro/Documents/Bayesian Statistics/Project/Cpp_code/GibbsSampler_updated.cpp")
sourceCpp("/Users/macbookpro/Documents/Bayesian Statistics/Project/Cpp_code/mcmcPotts.cpp")
```


##loading data

```{r}
D = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-IM-Step1-Step2-Step4-Step5-101.txt")
D0 = D
D0[is.na(D0)] = 0

pixels = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-XYCoordinates-Step1-Step2-Step4-Step5-101.txt")
colnames(D0) = substr(colnames(D0),1,4)
colnames(pixels) = c("x","y")

max_number_of_pixels = c(max(pixels[,1]),max(pixels[,2]))
```

## convenional pca on data

```{r}
pca = princomp(D0)
plot(pca)
summary(pca)
```

we start with using two components since this gives us 97% of explained variance

```{r}
df <- data.frame("pca1" = pca$scores[,1], "pca2" = pca$scores[,2], "x" = pixels$x, "y" = pixels$y)
data = rbind(pca$scores[,1],pca$scores[,2])
P1 = ggplot(data = df, aes(x = pca1,after_stat(density))) + geom_histogram(bins = 200)+ geom_density(color = "red",linewidth = 1)
P2 = ggplot(data = df, aes(x = pca2,after_stat(density))) + geom_histogram(bins = 200)+ geom_density(color = "blue",linewidth = 1)
P1 + P2
```

```{r}
PCA1 = ggplot(df) + geom_tile(aes(x=x,y=y,fill = pca1)) + scale_fill_viridis_c(option = "H") + theme_void() +
 theme(legend.position = "bottom") 

PCA2 = ggplot(df) + geom_tile(aes(x=x,y=y,fill = pca2)) + scale_fill_viridis_c(option = "H") + theme_void() +
 theme(legend.position = "bottom") 

PCA1 + PCA2
```

```{r}
ggplot(data = df,aes(x = pca1,y = pca2)) + geom_point(alpha = 0.1) + geom_density_2d(bins = 20) + geom_density2d_filled(alpha = 0.1)
```

most of the data is in the right part of the graph

```{r,warning=FALSE}
P1 = ggplot(data = df,aes(x = pca1,y = pca2)) + geom_point(alpha = 0.3) +
  xlim(10,20) + ylim(-3,3) + geom_density_2d() + geom_density2d_filled(alpha = 0.25)
P2 = ggplot(data = df,aes(x = pca1,y = pca2)) + geom_point(alpha = 0.3) +
  xlim(17,19.5) + ylim(0,2) + geom_density_2d()
P2 + P1
```

# K = 3

## GMM

we start performing a 2d GMM with non informative priors on the data

```{r}
k = 3
d = 2
B0 = diag(100,d,d)
V0 = diag(10,d,d)
n0 = 10

priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
priors$lambda <- rep(1,k)
```

```{r}
iter = 10000
burnin = 5000
```

```{r}
resGMM <- GibbsGMM(data,iter,burnin,priors)
```


```{r}
allocation = rep(0,dim(pixels)[1])
for (i in 1:dim(pixels)[1]){
  allocation[i] = which.max(resGMM$alloc[i,])
}
df$alloc = allocation
```

```{r}
ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = as.factor(alloc))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")
#ggsave("GMM_k3.png")
```

let's plot the different clusters

```{r}
ggplot(data = df,aes(x = pca1,y = pca2,color = as.factor(alloc))) + geom_point(alpha = 0.3) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + labs(fill = "cluster")
```

let us plot the chains

```{r}
muchain = data.frame("mu" = c(resGMM$mu[1,1,], resGMM$mu[2,1,],resGMM$mu[1,2,], resGMM$mu[2,2,],
                     resGMM$mu[1,3,], resGMM$mu[2,3,]),"index" = rep(seq(1,iter),d*k),
                     "cluster" = c(rep(1,2*iter),rep(2,2*iter),rep(3,2*iter)), "comp" = rep(c(rep(1,iter),rep(2,iter))),k)

ggplot(data = muchain,aes(x = index,y = mu,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") + facet_wrap(~comp,scales = "free") + theme(legend.position = "bottom") + labs(color = "cluster")
```

let us calculate the MC estimates of the means

```{r}
mu1 = rowMeans(resGMM$mu[,1,(burnin +1):iter])
mu2 = rowMeans(resGMM$mu[,2,(burnin +1):iter])
mu3 = rowMeans(resGMM$mu[,3,(burnin +1):iter])

mudf = data.frame("x" = c(mu1[1],mu2[1],mu3[1]),"y" = c(mu1[2],mu2[2],mu3[2]),"alloc" = c(1,2,3))
```

and for the covariance matrices

```{r}

sigma1chain = array(dim = c(d,d,iter))
sigma2chain = array(dim = c(d,d,iter))
sigma3chain = array(dim = c(d,d,iter))
for(i in 1:iter){
  sigma1chain[,,i] = resGMM$sigma[i][[1]][,,1]
  sigma2chain[,,i] = resGMM$sigma[i][[1]][,,2]
  sigma3chain[,,i] = resGMM$sigma[i][[1]][,,3]
}

sigma1 = apply(sigma1chain, c(1,2), mean)
sigma2 = apply(sigma2chain, c(1,2), mean)
sigma3 = apply(sigma3chain, c(1,2), mean)

sigmachain = data.frame("sigma" = c(sigma1chain[1,1,], sigma1chain[2,1,],sigma1chain[1,2,], sigma1chain[2,2,],
                     sigma2chain[1,1,], sigma2chain[2,1,],sigma2chain[1,2,], sigma2chain[2,2,],sigma3chain[1,1,],
                     sigma3chain[2,1,],sigma3chain[1,2,], sigma3chain[2,2,]),"index" = rep(seq(1,iter),d*d*k),
                     "cluster" = c(rep(1,d*d*iter),rep(2,d*d*iter),rep(3,d*d*iter)), "comp1" = rep(c(rep(1,iter),rep(2,iter)),d*k),
                     "comp2" = rep(c(rep(1,d*iter),rep(2,d*iter)),k), "comp" =
                       rep(c(rep(11,iter),rep(12,iter),rep(21,iter),rep(22,iter)),k))

```

```{r}
ggplot(data = sigmachain,aes(x = index,y = sigma,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + facet_wrap(~comp,scales = "free") + labs(color = "cluster") 
```

wee still need to plot the covariance autocorr functions

```{r}
wchain <- data.frame("w" = c(resGMM$w[,1],resGMM$w[,2],resGMM$w[,3]), "index" = rep(seq(1,iter),k), 
                     "cluster" =c(rep(1,iter),rep(2,iter),rep(3,iter)))

ggplot(data = wchain,aes(x = index,y = w,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H")  +
  theme(legend.position = "bottom") + labs(color = "cluster")
```

```{r}
w = colMeans(resGMM$w[(burnin+1):iter,])
```

let's investigate the results:

```{r}
ggplot(data = df,aes(x = pca1,y = pca2,color = as.factor(alloc))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + geom_point(data = mudf,aes(x,y),color = "red",size = 2) + facet_wrap(~as.factor(alloc),scales = "free") + labs(color = "cluster") 
```

```{r}
G1 = ggplot(data = df,aes(x = pca1,color = as.factor(alloc), fill = as.factor(alloc),after_stat(density))) + geom_histogram(bins = 100,alpha = 0,linewidth = 0.7) + geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G2 = ggplot(data = df,aes(x = pca2,color = as.factor(alloc), fill = as.factor(alloc),after_stat(density))) + geom_histogram(bins = 100,alpha = 0,linewidth = 0.7) + geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G1 + G2
```

the bimodal shape of one of the component is not wanted, we will try to increase the k

## Potts Model

we need to know the geometry, we do it through the mask 

```{r}
mask <-matrix(0,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  mask[pixels[i,1],pixels[i,2]] = 1
}

neigh <- getNeighbors(mask = mask, c(2,2,0,0))
block <- getBlocks(mask = mask, 2)
```

## Fixing the beta to betacritic

```{r}
priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
```

```{r}
resultsGibbs <- GibbsPotts(data,betacritic,initmu,initsigma,neigh,block,priors,iter,burnin)
```

```{r}
allocationgibbs = rep(0,dim(pixels)[1])
for (i in 1:dim(pixels)[1]){
  allocationgibbs[i] = which.max(resultsGibbs$alloc[i,])
  # just to match the labels to above
  if(allocationgibbs[i]== 1){
    allocationgibbs[i] = 3
  }
  else if(allocationgibbs[i]== 3){
    allocationgibbs[i] = 1
  }
}
df$allocgibbs= allocationgibbs
```

```{r}
ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocationgibbs))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")
#ggsave("gibbs_k3.png")
```

```{r}
ggplot(data = df,aes(x = pca1,y = pca2,color = as.factor(allocgibbs))) + geom_point(alpha = 0.5) + stat_ellipse(linewidth = 1) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + labs(color = "cluster")
#ggsave("gibbs_k3satter.png")
```

```{r}
diff_gmm_gibbs = rep(0,dim(pixels)[1])
for (i in 1:dim(pixels)[1]){
  if(df$alloc[i]!=df$allocgibbs[i]){
    diff_gmm_mcmc[i] = 1
  }
}
df$diff_gmm_gibbs = diff_gmm_gibbs
```

```{r}
ggplot(data = df,aes(x = pca1,y = pca2,color = as.factor(diff_gmm_gibbs))) + geom_point(alpha = 0.5) +  scale_colour_viridis_d(option = "plasma") + theme(legend.position = "bottom") + labs(color = "differently classified")
#ggsave("diff_gmm_mcmc_k3b.png")
```

```{r}
G1 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(diff_gmm_gibbs))) + scale_fill_viridis_d(option = "plasma") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "difference GMM Potts")
G2 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(alloc))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G3 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocationgibbs))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G2 + G1 + G3
#ggsave("diff_gmm_mcmc_k3.png")
```
the difference is negligeble

### Pseudolikelihood

```{r}
betacritic = log(1 + sqrt(k))
initmu = matrix(0,d,k)
initsigma = array(n0*V0,dim = c(d,d,k))

priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
priors$beta <- c(0,2*betacritic)

mh <- list(bandwidth=0.1)
```

```{r}
resultsMCMC <- MCMCPotts(data,neigh,block,iter,burnin,priors,mh)
```

```{r}
allocationmcmc = rep(0,dim(pixels)[1])
for (i in 1:dim(pixels)[1]){
  allocationmcmc[i] = which.max(resultsMCMC$alloc[i,])
  # just to match the labels to above
  if(allocationmcmc[i]== 3){
    allocationmcmc[i] = 1
  }
  else if(allocationmcmc[i]== 1){
    allocationmcmc[i] = 2
  }
  else if(allocationmcmc[i]== 2){
    allocationmcmc[i] = 3
  }
}

df$allocmcmc = allocationmcmc
```

```{r}
ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocmcmc))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")
#ggsave("mcmc_k3.png")
```

let's plot the different clusters

```{r}
ggplot(data = df,aes(x = pca1,y = pca2,color = as.factor(allocmcmc))) + geom_point(alpha = 0.5) + stat_ellipse(linewidth = 1) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + labs(color = "cluster")
#ggsave("mcmc_k3_scatter.png")
```

```{r}
muchain = data.frame("mu" = c(resultsMCMC$mu[1,1,], resultsMCMC$mu[2,1,],resultsMCMC$mu[1,2,], resultsMCMC$mu[2,2,],
                     resultsMCMC$mu[1,3,], resultsMCMC$mu[2,3,]),"index" = rep(seq(1,iter),d*k),
                     "cluster" = c(rep(1,2*iter),rep(2,2*iter),rep(3,2*iter)), "comp" = rep(c(rep(1,iter),rep(2,iter))),k)

ggplot(data = muchain,aes(x = index,y = mu,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") + facet_wrap(~comp,scales = "free") + theme(legend.position = "bottom") + labs(color = "cluster")

```

let us calculate the MC estimates of the means

```{r}
mu1 = rowMeans(resultsMCMC$mu[,1,(burnin +1):iter])
mu2 = rowMeans(resultsMCMC$mu[,2,(burnin +1):iter])
mu3 = rowMeans(resultsMCMC$mu[,3,(burnin +1):iter])

mudf = data.frame("x" = c(mu1[1],mu2[1],mu3[1]),"y" = c(mu1[2],mu2[2],mu3[2]),"alloc" = c(1,2,3))
```

and for the covariance matrices

```{r}

sigma1chain = array(dim = c(d,d,iter))
sigma2chain = array(dim = c(d,d,iter))
sigma3chain = array(dim = c(d,d,iter))
for(i in 1:iter){
  sigma1chain[,,i] = resultsMCMC$sigma[i][[1]][,,1]
  sigma2chain[,,i] = resultsMCMC$sigma[i][[1]][,,2]
  sigma3chain[,,i] = resultsMCMC$sigma[i][[1]][,,3]
}

sigma1 = apply(sigma1chain, c(1,2), mean)
sigma2 = apply(sigma2chain, c(1,2), mean)
sigma3 = apply(sigma3chain, c(1,2), mean)

sigmachain = data.frame("sigma" = c(sigma1chain[1,1,], sigma1chain[2,1,],sigma1chain[1,2,], sigma1chain[2,2,],
                     sigma2chain[1,1,], sigma2chain[2,1,],sigma2chain[1,2,], sigma2chain[2,2,],sigma3chain[1,1,],
                     sigma3chain[2,1,],sigma3chain[1,2,], sigma3chain[2,2,]),"index" = rep(seq(1,iter),d*d*k),
                     "cluster" = c(rep(1,d*d*iter),rep(2,d*d*iter),rep(3,d*d*iter)), "comp1" = rep(c(rep(1,iter),rep(2,iter)),d*k),
                     "comp2" = rep(c(rep(1,d*iter),rep(2,d*iter)),k), "comp" =
                       rep(c(rep(11,iter),rep(12,iter),rep(21,iter),rep(22,iter)),k))

```

```{r}
ggplot(data = sigmachain,aes(x = index,y = sigma,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + facet_wrap(~comp,scales = "free") + labs(color = "cluster") 
```

```{r}
sumchain <- data.frame("sum" = resultsMCMC$sum[,1], "index" = seq(1,iter))
betachain <- data.frame("beta" = resultsMCMC$beta[,1], "index" = seq(1,iter))

G1 = ggplot(data = sumchain,aes(x = index,y = sum)) + geom_line() 
G2 = ggplot(data = betachain,aes(x = index,y = beta)) + geom_line() + geom_hline(yintercept = betacritic,color = "red",)
G1 + G2
```

here beta went above beta critic, but we were able to correctly identify the beta and the clusters, all the chains are relatively well behaved

let us look at the differences between the potts model and the GMM

```{r}
diff_gmm_mcmc = rep(0,dim(pixels)[1])
for (i in 1:dim(pixels)[1]){
  if(df$alloc[i]!=df$allocmcmc[i]){
    diff_gmm_mcmc[i] = 1
  }
}
df$diff_gmm_mcmc = diff_gmm_mcmc
```

```{r}
ggplot(data = df,aes(x = pca1,y = pca2,color = as.factor(diff_gmm_mcmc))) + geom_point(alpha = 0.5) +  scale_colour_viridis_d(option = "plasma") + theme(legend.position = "bottom") + labs(color = "differently classified")
#ggsave("diff_gmm_mcmc_k3b.png")
```

```{r}
G1 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(diff_gmm_mcmc))) + scale_fill_viridis_d(option = "plasma") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "difference GMM Potts")
G2 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(alloc))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G3 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocmcmc))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G2 + G1 + G3
ggsave("diff_gmm_mcmc_k3.png")
```

the Potts model tend to make the clusters more unoform, there is not much difference in the shapes, just the extent of the edges is affected 

let us compare the two different potts model, the one with fixed beta and the one with the sampled beta

```{r}
diff_gibbs_mcmc = rep(0,dim(pixels)[1])
for (i in 1:dim(pixels)[1]){
  if(df$allocgibbs[i]!=df$allocmcmc[i]){
    diff_gibbs_mcmc[i] = 1
  }
}
df$diff_gibbs_mcmc = diff_gibbs_mcmc
```

```{r}
G1 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(diff_gibbs_mcmc))) + scale_fill_viridis_d(option = "plasma") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "difference GMM Potts")
G2 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocgibbs))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G3 = ggplot(df)+
  geom_tile(aes(x=x,y=y,fill = factor(allocmcmc))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G2 + G1 + G3
#ggsave("diff_gibbs_mcmc_k3.png")
```

the difference is negligible in this case

# K = 6

## GMM

we start performing a 2d GMM with non informative priors on the data

```{r}
k = 6
d = 2
B0 = diag(100,d,d)
V0 = diag(10,d,d)
n0 = 10

priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
priors$lambda <- rep(1,k)
```

```{r}
resGMM <- GibbsGMM(data,iter,burnin,priors)
```

```{r}
df2 <- data.frame("pca1" = pca$scores[,1], "pca2" = pca$scores[,2], "x" = pixels$x, "y" = pixels$y)
```

```{r}
allocation = rep(0,dim(pixels)[1])
for (i in 1:dim(pixels)[1]){
  allocation[i] = which.max(resGMM$alloc[i,])
}
df2$alloc = allocation
```

```{r}
ggplot(df2)+
  geom_tile(aes(x=x,y=y,fill = as.factor(alloc))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")
ggsave("gmm_k6cluster.png")
```

```{r}
print(table(df2$alloc))
```

let's plot the different clusters

```{r}
ggplot(data = df2,aes(x = pca1,y = pca2,color = as.factor(alloc))) + geom_point(alpha = 0.5) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + labs(color = "cluster")
ggsave("gmm_k6scattera.png")
```


```{r}
ggplot(data = df2,aes(x = pca1,y = pca2,color = as.factor(alloc))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(alloc),scales = "free") + labs(color = "cluster") 
ggsave("gmm_k6scatterb.png")
```

```{r}
G1 = ggplot(data = df2,aes(x = pca1,color = as.factor(alloc), fill = as.factor(alloc),after_stat(density))) + geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G2 = ggplot(data = df2,aes(x = pca2,color = as.factor(alloc), fill = as.factor(alloc),after_stat(density))) + geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G1 + G2
ggsave("gmm_k6marginal.png")
```

let us plot the chains

```{r}
muchain = data.frame("mu" = c(resGMM$mu[1,1,], resGMM$mu[2,1,],resGMM$mu[1,2,], resGMM$mu[2,2,],
                     resGMM$mu[1,3,], resGMM$mu[2,3,],resGMM$mu[1,4,], resGMM$mu[2,4,],
                     resGMM$mu[1,5,], resGMM$mu[2,5,],resGMM$mu[1,6,], resGMM$mu[2,6,]),"index" = rep(seq(1,iter),d*k),
                     "cluster" = c(rep(1,2*iter),rep(2,2*iter),rep(3,2*iter),rep(4,2*iter),rep(5,2*iter),rep(6,2*iter)), "comp" = rep(c(rep(1,iter),rep(2,iter))),k)

ggplot(data = muchain,aes(x = index,y = mu,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") + facet_wrap(~comp,scales = "free") + theme(legend.position = "bottom") + labs(color = "cluster")
ggsave("gmm_k6muchain.png")
```

let us calculate the MC estimates of the means

```{r}
mu1 = rowMeans(resGMM$mu[,1,(burnin +1):iter])
mu2 = rowMeans(resGMM$mu[,2,(burnin +1):iter])
mu3 = rowMeans(resGMM$mu[,3,(burnin +1):iter])
mu4 = rowMeans(resGMM$mu[,4,(burnin +1):iter])
mu5 = rowMeans(resGMM$mu[,5,(burnin +1):iter])
mu6 = rowMeans(resGMM$mu[,6,(burnin +1):iter])

mudf = data.frame("x" = c(mu1[1],mu2[1],mu3[1],mu4[1],mu5[1],mu6[1]),"y" = c(mu1[2],mu2[2],mu3[2],mu4[2],mu5[2],mu6[2]),"alloc" = c(1,2,3,4,5,6))
```

and for the covariance matrices

```{r}

sigma1chain = array(dim = c(d,d,iter))
sigma2chain = array(dim = c(d,d,iter))
sigma3chain = array(dim = c(d,d,iter))
sigma4chain = array(dim = c(d,d,iter))
sigma5chain = array(dim = c(d,d,iter))
sigma6chain = array(dim = c(d,d,iter))

for(i in 1:iter){
  sigma1chain[,,i] = resGMM$sigma[i][[1]][,,1]
  sigma2chain[,,i] = resGMM$sigma[i][[1]][,,2]
  sigma3chain[,,i] = resGMM$sigma[i][[1]][,,3]
  sigma4chain[,,i] = resGMM$sigma[i][[1]][,,4]
  sigma5chain[,,i] = resGMM$sigma[i][[1]][,,5]
  sigma6chain[,,i] = resGMM$sigma[i][[1]][,,6]
}

sigma1 = apply(sigma1chain, c(1,2), mean)
sigma2 = apply(sigma2chain, c(1,2), mean)
sigma3 = apply(sigma3chain, c(1,2), mean)
sigma4 = apply(sigma4chain, c(1,2), mean)
sigma5 = apply(sigma5chain, c(1,2), mean)
sigma6 = apply(sigma5chain, c(1,2), mean)

sigmachain = data.frame("sigma" = c(sigma1chain[1,1,], sigma1chain[2,1,],sigma1chain[1,2,], sigma1chain[2,2,],
                     sigma2chain[1,1,], sigma2chain[2,1,],sigma2chain[1,2,], sigma2chain[2,2,],sigma3chain[1,1,],
                     sigma3chain[2,1,],sigma3chain[1,2,], sigma3chain[2,2,], sigma4chain[1,1,], sigma4chain[2,1,],sigma4chain[1,2,],
                     sigma4chain[2,2,],sigma5chain[1,1,],sigma5chain[2,1,],sigma5chain[1,2,], sigma5chain[2,2,],
                     sigma6chain[1,1,],sigma6chain[2,1,],sigma6chain[1,2,], sigma6chain[2,2,] ),
                     "index" = rep(seq(1,iter),d*d*k),"cluster" = 
                       c(rep(1,d*d*iter),rep(2,d*d*iter),rep(3,d*d*iter),rep(4,d*d*iter),rep(5,d*d*iter),rep(6,d*d*iter)),
                     "comp1" = rep(c(rep(1,iter),rep(2,iter)),d*k),
                     "comp2" = rep(c(rep(1,d*iter),rep(2,d*iter)),k), "comp" =
                       rep(c(rep(11,iter),rep(12,iter),rep(21,iter),rep(22,iter)),k))

```

```{r}
ggplot(data = sigmachain,aes(x = index,y = sigma,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + facet_wrap(~comp,scales = "free") + labs(color = "cluster") 
ggsave("gmm_k6sigmachain.png")
```

wee still need to plot the covariance autocorr functions

```{r}
wchain <- data.frame("w" = c(resGMM$w[,1],resGMM$w[,2],resGMM$w[,3],resGMM$w[,4],resGMM$w[,5],resGMM$w[,6]), "index" = rep(seq(1,iter),k), 
                     "cluster" =c(rep(1,iter),rep(2,iter),rep(3,iter),rep(4,iter),rep(5,iter),rep(6,iter)))

ggplot(data = wchain,aes(x = index,y = w,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H")  +
  theme(legend.position = "bottom") + labs(color = "cluster")
ggsave("gmm_k6wchain.png")
```

```{r}
w = colMeans(resGMM$w[(burnin+1):iter,])
```

## Potts Model

### Pseudolikelihood

```{r}
betacritic = log(1 + sqrt(k))
initmu = matrix(0,d,k)
initsigma = array(n0*V0,dim = c(d,d,k))

priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
priors$beta <- c(0,2*betacritic)

mh <- list(bandwidth=0.1)
```

```{r}
resultsMCMC <- MCMCPotts(data,neigh,block,iter,burnin,priors,mh)
```

```{r}
allocationmcmc = rep(0,dim(pixels)[1])
for (i in 1:dim(pixels)[1]){
  allocationmcmc[i] = which.max(resultsMCMC$alloc[i,])
  # just to match the labels to above
  if(allocationmcmc[i]== 5){
    allocationmcmc[i] = 2
  }
  else if(allocationmcmc[i]== 2){
    allocationmcmc[i] = 1
  }
  else if(allocationmcmc[i]== 1){
    allocationmcmc[i] = 3
  }
  else if(allocationmcmc[i]== 3){
    allocationmcmc[i] = 5
  }
}

df2$allocmcmc = allocationmcmc
```

```{r}
ggplot(df2)+
  geom_tile(aes(x=x,y=y,fill = factor(allocmcmc))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")
ggsave("mcmc_k6cluster.png")
```

let's plot the different clusters

```{r}
ggplot(data = df2,aes(x = pca1,y = pca2,color = as.factor(allocmcmc))) + geom_point(alpha = 0.5) + stat_ellipse(linewidth = 1) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + labs(color = "cluster")
ggsave("mcmc_k6scatter.png")
```



```{r}
ggplot(data = df2,aes(x = pca1,y = pca2,color = as.factor(allocmcmc))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(allocmcmc),scales = "free") + labs(color = "cluster") 

ggsave("mcmc_k6scatterb.png")
```

```{r}
G1 = ggplot(data = df2,aes(x = pca1,color = as.factor(allocmcmc), fill = as.factor(allocmcmc),after_stat(density))) + geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G2 = ggplot(data = df2,aes(x = pca2,color = as.factor(allocmcmc), fill = as.factor(allocmcmc),after_stat(density))) + geom_density(linewidth = 1.2)  + scale_colour_viridis_d(option = "H") + scale_fill_viridis_d(option = "H",alpha = 0.3) + theme(legend.position = "bottom") + labs(color = "cluster", fill = "cluster")

G1 + G2
ggsave("mcmc_k6marginal.png")
```

```{r}
muchain = data.frame("mu" = c(resultsMCMC$mu[1,1,], resultsMCMC$mu[2,1,],resultsMCMC$mu[1,2,], resultsMCMC$mu[2,2,],
                     resultsMCMC$mu[1,3,], resultsMCMC$mu[2,3,],resultsMCMC$mu[1,4,], resultsMCMC$mu[2,4,],
                     resultsMCMC$mu[1,5,], resultsMCMC$mu[2,5,],resultsMCMC$mu[1,6,], resultsMCMC$mu[2,6,]),"index" = rep(seq(1,iter),d*k),
                     "cluster" = c(rep(1,2*iter),rep(2,2*iter),rep(3,2*iter),rep(4,2*iter),rep(5,2*iter),rep(6,2*iter)), "comp" = rep(c(rep(1,iter),rep(2,iter))),k)

ggplot(data = muchain,aes(x = index,y = mu,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") + facet_wrap(~comp,scales = "free") + theme(legend.position = "bottom") + labs(color = "cluster")
ggsave("mcmc_k6muchain.png")
```

```{r}

sigma1chain = array(dim = c(d,d,iter))
sigma2chain = array(dim = c(d,d,iter))
sigma3chain = array(dim = c(d,d,iter))
sigma4chain = array(dim = c(d,d,iter))
sigma5chain = array(dim = c(d,d,iter))
sigma6chain = array(dim = c(d,d,iter))

for(i in 1:iter){
  sigma1chain[,,i] = resultsMCMC$sigma[i][[1]][,,1]
  sigma2chain[,,i] = resultsMCMC$sigma[i][[1]][,,2]
  sigma3chain[,,i] = resultsMCMC$sigma[i][[1]][,,3]
  sigma4chain[,,i] = resultsMCMC$sigma[i][[1]][,,4]
  sigma5chain[,,i] = resultsMCMC$sigma[i][[1]][,,5]
  sigma6chain[,,i] = resultsMCMC$sigma[i][[1]][,,6]
}

sigma1 = apply(sigma1chain, c(1,2), mean)
sigma2 = apply(sigma2chain, c(1,2), mean)
sigma3 = apply(sigma3chain, c(1,2), mean)
sigma4 = apply(sigma4chain, c(1,2), mean)
sigma5 = apply(sigma5chain, c(1,2), mean)
sigma5 = apply(sigma6chain, c(1,2), mean)

sigmachain = data.frame("sigma" = c(sigma1chain[1,1,], sigma1chain[2,1,],sigma1chain[1,2,], sigma1chain[2,2,],
                     sigma2chain[1,1,], sigma2chain[2,1,],sigma2chain[1,2,], sigma2chain[2,2,],sigma3chain[1,1,],
                     sigma3chain[2,1,],sigma3chain[1,2,], sigma3chain[2,2,], sigma4chain[1,1,], sigma4chain[2,1,],sigma4chain[1,2,],
                     sigma4chain[2,2,],sigma5chain[1,1,],sigma5chain[2,1,],sigma5chain[1,2,], sigma5chain[2,2,],
                     sigma6chain[1,1,],sigma6chain[2,1,],sigma6chain[1,2,], sigma6chain[2,2,] ),
                     "index" = rep(seq(1,iter),d*d*k),"cluster" = 
                       c(rep(1,d*d*iter),rep(2,d*d*iter),rep(3,d*d*iter),rep(4,d*d*iter),rep(5,d*d*iter),rep(6,d*d*iter)),
                     "comp1" = rep(c(rep(1,iter),rep(2,iter)),d*k),
                     "comp2" = rep(c(rep(1,d*iter),rep(2,d*iter)),k), "comp" =
                       rep(c(rep(11,iter),rep(12,iter),rep(21,iter),rep(22,iter)),k))

```

```{r}
ggplot(data = sigmachain,aes(x = index,y = sigma,color = as.factor(cluster))) + geom_line() + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + facet_wrap(~comp,scales = "free") + labs(color = "cluster") 
ggsave("mcmc_k6sigmachain.png")
```

```{r}
sumchain <- data.frame("sum" = resultsMCMC$sum[,1], "index" = seq(1,iter))
betachain <- data.frame("beta" = resultsMCMC$beta[,1], "index" = seq(1,iter))

G1 = ggplot(data = sumchain,aes(x = index,y = sum)) + geom_line() 
G2 = ggplot(data = betachain,aes(x = index,y = beta)) + geom_line() + geom_hline(yintercept = betacritic,color = "red",)
G1 + G2
```

here beta went above beta critic, but we were able to correctly identify the beta and the clusters, all the chains are relatively well behaved

```{r}

G2 = ggplot(df2)+
  geom_tile(aes(x=x,y=y,fill = factor(alloc))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G3 = ggplot(df2)+
  geom_tile(aes(x=x,y=y,fill = factor(allocmcmc))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G2 + G3
ggsave("mcmc_vs_gmmk6.png")
```

we have a considerable difference in this case

## Fixing the beta to betacritic

```{r}
priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
```

```{r}
resultsGibbs <- GibbsPotts(data,betacritic,initmu,initsigma,neigh,block,priors,iter,burnin)
```

```{r}
allocationgibbs = rep(0,dim(pixels)[1])
for (i in 1:dim(pixels)[1]){
  allocationgibbs[i] = which.max(resultsGibbs$alloc[i,])
  # just to match the labels to above
  if(allocationgibbs[i]== 3){
    allocationgibbs[i] = 1
  }
  else if(allocationgibbs[i]== 5){
    allocationgibbs[i] = 2
  }
  else if(allocationgibbs[i]== 1){
    allocationgibbs[i] = 5
  }
  else if(allocationgibbs[i]== 2){
    allocationgibbs[i] = 3
  }
}
df2$allocgibbs= allocationgibbs
```

```{r}
ggplot(df2)+
  geom_tile(aes(x=x,y=y,fill = factor(allocationgibbs))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")
ggsave("gibbs_k6.png")
```

```{r}
ggplot(data = df2,aes(x = pca1,y = pca2,color = as.factor(allocgibbs))) + geom_point(alpha = 0.5) + stat_ellipse(linewidth = 1) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + labs(color = "cluster")
ggsave("gibbs_k6scatter.png")
```

```{r}
ggplot(data = df2,aes(x = pca1,y = pca2,color = as.factor(allocgibbs))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(allocgibbs),scales = "free") + labs(color = "cluster") 
ggsave("gibbs_k6scatterb.png")
```

let us compare the two different potts model, the one with fixed beta and the one with the sampled beta

```{r}
diff_gibbs_mcmc = rep(0,dim(pixels)[1])
for (i in 1:dim(pixels)[1]){
  if(df2$allocgibbs[i]!=df2$allocmcmc[i]){
    diff_gibbs_mcmc[i] = 1
  }
}
df2$diff_gibbs_mcmc = diff_gibbs_mcmc
```

```{r}
G1 = ggplot(df2)+
  geom_tile(aes(x=x,y=y,fill = factor(diff_gibbs_mcmc))) + scale_fill_viridis_d(option = "plasma") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "difference GMM Potts")
G2 = ggplot(df2)+
  geom_tile(aes(x=x,y=y,fill = factor(allocgibbs))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G3 = ggplot(df2)+
  geom_tile(aes(x=x,y=y,fill = factor(allocmcmc))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "cluster")

G2 + G1 + G3
ggsave("diff_mmc_gibbs_k6.png")
```

the difference is quite significant


```{r}
G1 = ggplot(df2)+
  geom_tile(aes(x=x,y=y,fill = factor(alloc))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "GMM")
G2 = ggplot(df2)+
  geom_tile(aes(x=x,y=y,fill = factor(allocgibbs))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "Gibbs")

G3 = ggplot(df2)+
  geom_tile(aes(x=x,y=y,fill = factor(allocmcmc))) + scale_fill_viridis_d(option = "H") +
  theme_void() + theme(legend.position = "bottom") + labs(fill = "MCMC")

G1 + G2 + G3
ggsave("diff_gmm_mmc_gibbs_k6.png")
```

```{r}
G1 = ggplot(data = df2,aes(x = pca1,y = pca2,color = as.factor(alloc))) + geom_point(alpha = 0.5) + stat_ellipse(linewidth = 1) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + labs(color = "GMM")
G1a <- ggMarginal(G1,type = "density",groupColour = TRUE, groupFill = TRUE)
G2 = ggplot(data = df2,aes(x = pca1,y = pca2,color = as.factor(allocgibbs))) + geom_point(alpha = 0.5) + stat_ellipse(linewidth = 1) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + labs(color = "Gibbs")
G2a <- ggMarginal(G2,type = "density",groupColour = TRUE, groupFill = TRUE)
G3 = ggplot(data = df2,aes(x = pca1,y = pca2,color = as.factor(allocmcmc))) + geom_point(alpha = 0.5) + stat_ellipse(linewidth = 1) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") + labs(color = "MCMC")
G3a <- ggMarginal(G3,type = "density",groupColour = TRUE, groupFill = TRUE)

G1 + G2 + G3
ggsave("diff_gmm_mmc_gibbs_k6scatter.png")
```

```{r}
G1 = ggplot(data = df2,aes(x = pca1,y = pca2,color = as.factor(alloc))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(alloc),scales = "free") + labs(color = "cluster") 

G2 = ggplot(data = df2,aes(x = pca1,y = pca2,color = as.factor(allocgibbs))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(allocgibbs),scales = "free") + labs(color = "cluster")

G3 = ggplot(data = df2,aes(x = pca1,y = pca2,color = as.factor(allocmcmc))) + geom_point(alpha = 0.2) + scale_colour_viridis_d(option = "H") +
  theme(legend.position = "bottom") +  stat_ellipse(linewidth = 1) + facet_wrap(~as.factor(allocmcmc),scales = "free") + labs(color = "cluster")

G1 / G2 / G3
ggsave("diff_gmm_mmc_gibbs_k6scatterb.png")
```

```{r}
G1a
ggsave("gmm_k6_full.png",width = 20,units = "cm")
```

```{r}
G2a
ggsave("gibbs_k6_full.png",width = 20,units = "cm")
```

```{r}
G3a
ggsave("mcmc_k6_full.png",width = 20,units = "cm")
```

