---
title: "multidimensional potts model lipidi"
author: Simone Colombara, Alessia Cotroneo, Francesco De Caro, Riccardo Morandi, Chiara   Schembri,
  Alfredo Zapiola
date: "2022-12-16"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rayshader)
library(patchwork)
library(viridis)
library(plot.matrix)
library(bayesImageS)
library(stats)
library(Rcpp)
library(RcppArmadillo)
library(coda)
library(MASS)
```


```{r, warning=FALSE}
sourceCpp("/Users/macbookpro/Documents/Bayesian Statistics/Project/Cpp_code/GibbsSampler_updated.cpp")
sourceCpp("/Users/macbookpro/Documents/Bayesian Statistics/Project/Cpp_code/mcmcPotts.cpp")
```


##loading data

```{r}
D = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-IM-Step1-Step2-Step4-Step5-101.txt")
D0 = D
D0[is.na(D0)] = 0

pixels = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-XYCoordinates-Step1-Step2-Step4-Step5-101.txt")
colnames(D0) = substr(colnames(D0),1,4)
colnames(pixels) = c("x","y")

max_number_of_pixels = c(max(pixels[,1]),max(pixels[,2]))
```

## convenional pca on data

```{r}
pca = princomp(D0)
plot(pca)
summary(pca)
```

# d = 2

we start with using two components since this gives us 97% of explained variance

```{r}
pcascore1vec = pca$scores[,1]
pcascore2vec = pca$scores[,2]
data = rbind(pcascore1vec,pcascore2vec)
```

```{r}
plot(pcascore1vec,pcascore2vec)
est12 = kde2d(pcascore1vec, pcascore2vec,  n = 100)
contour(est12,nlevels = 50,add = TRUE,col = hcl.colors(50, "Spectral"))
```

### single component plots

```{r}
comp1 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
comp2 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  comp1[pixels[i,1],pixels[i,2]] = pcascore1vec[i]
  comp2[pixels[i,1],pixels[i,2]] = pcascore2vec[i]
}
par(mfrow=c(1,2))
plot(comp1, border=NA,asp = TRUE,col= turbo,axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
plot(comp2, border=NA,asp = TRUE,col= turbo,axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```

```{r}
mask <-matrix(0,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  mask[pixels[i,1],pixels[i,2]] = 1
}

neigh <- getNeighbors(mask = mask, c(2,2,0,0))
block <- getBlocks(mask = mask, 2)
```

## K = 3

we start with non informative priors for the parameters

```{r}
k = 3
betacritic = log(1 + sqrt(k))
d = 2
B0 = diag(100,d,d)
V0 = diag(10,d,d)
n0 = 10
initmu = matrix(0,d,k)
initsigma = array(n0*V0,dim = c(d,d,k))

priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)

```

```{r}
iter = 10000
burnin = 5000
results <- GibbsPotts(data,betacritic,initmu,initsigma,neigh,block,priors,iter,burnin)
```

```{r}
clustering <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering[pixels[i,1],pixels[i,2]] = which.max(results$alloc[i,])
}
print(table(clustering))
```

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering, border=NA,asp = TRUE,col= turbo(k),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```

```{r}
plot(pcascore1vec,pcascore2vec,col = clustering[pixels[,1],pixels[,2]])
```

```{r}
mu1chain = mcmc(t(results$mu[,1,]))
mu2chain = mcmc(t(results$mu[,2,]))
mu3chain = mcmc(t(results$mu[,3,]))
```

```{r}
par(mfrow = c(d,d))
plot(mu1chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu1chain,auto.layout = FALSE)
plot(mu2chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu2chain,auto.layout = FALSE)
plot(mu3chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu3chain,auto.layout = FALSE)
```

```{r}
trace1 = rep(0,iter)
trace2 = rep(0,iter)
trace3 = rep(0,iter)

for( i in 1:iter){
  trace1[i] = sum(diag(results$sigmas[i][[1]][,,1]))
  trace2[i] = sum(diag(results$sigmas[i][[1]][,,2]))
  trace3[i] = sum(diag(results$sigmas[i][[1]][,,3]))
}

tr1chain = mcmc(trace1)
tr2chain = mcmc(trace2)
tr3chain = mcmc(trace3)

par(mfrow = c(3,2))
plot(tr1chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr1chain,auto.layout = FALSE)
plot(tr2chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr2chain,auto.layout = FALSE)
plot(tr3chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr3chain,auto.layout = FALSE)
```

## K = 6

we start with non informative priors for the parameters

```{r}
k = 6
betacritic = log(1 + sqrt(k))
d = 2
B0 = diag(100,d,d)
V0 = diag(10,d,d)
n0 = 10
initmu = matrix(0,d,k)
initsigma = array(n0*V0,dim = c(d,d,k))

priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
```

```{r}
iter = 10000
burnin = 5000
results <- GibbsPotts(data,betacritic,initmu,initsigma,neigh,block,priors,iter,burnin)
```

```{r}
clustering <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering[pixels[i,1],pixels[i,2]] = which.max(results$alloc[i,])
}
print(table(clustering))
```

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering, border=NA,asp = TRUE,col= turbo(k),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```

```{r}
plot(pcascore1vec,pcascore2vec,col = clustering[pixels[,1],pixels[,2]])
```

```{r}
mu1chain = mcmc(t(results$mu[,1,]))
mu2chain = mcmc(t(results$mu[,2,]))
mu3chain = mcmc(t(results$mu[,3,]))
mu4chain = mcmc(t(results$mu[,4,]))
mu5chain = mcmc(t(results$mu[,5,]))
mu6chain = mcmc(t(results$mu[,6,]))
```


```{r}
par(mfrow = c(d,d))
plot(mu1chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu1chain,auto.layout = FALSE)
plot(mu2chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu2chain,auto.layout = FALSE)
plot(mu3chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu3chain,auto.layout = FALSE)
plot(mu4chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu4chain,auto.layout = FALSE)
plot(mu5chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu5chain,auto.layout = FALSE)
plot(mu6chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu6chain,auto.layout = FALSE)
```

```{r}
trace1 = rep(0,iter)
trace2 = rep(0,iter)
trace3 = rep(0,iter)
trace4 = rep(0,iter)
trace5 = rep(0,iter)
trace6 = rep(0,iter)

for( i in 1:iter){
  trace1[i] = sum(diag(results$sigmas[i][[1]][,,1]))
  trace2[i] = sum(diag(results$sigmas[i][[1]][,,2]))
  trace3[i] = sum(diag(results$sigmas[i][[1]][,,3]))
  trace4[i] = sum(diag(results$sigmas[i][[1]][,,4]))
  trace5[i] = sum(diag(results$sigmas[i][[1]][,,5]))
  trace6[i] = sum(diag(results$sigmas[i][[1]][,,6]))
}

tr1chain = mcmc(trace1)
tr2chain = mcmc(trace2)
tr3chain = mcmc(trace3)
tr4chain = mcmc(trace4)
tr5chain = mcmc(trace5)
tr6chain = mcmc(trace6)

par(mfrow = c(2,2))
plot(tr1chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr1chain,auto.layout = FALSE)
plot(tr2chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr2chain,auto.layout = FALSE)
plot(tr3chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr3chain,auto.layout = FALSE)
plot(tr4chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr4chain,auto.layout = FALSE)
plot(tr5chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr5chain,auto.layout = FALSE)
plot(tr6chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr6chain,auto.layout = FALSE)
```

# bata approximated via pseudolikelihood

## K = 3

we start with non informative priors for the parameters

```{r}
k = 3
betacritic = log(1 + sqrt(k))
d = 2
B0 = diag(100,d,d)
V0 = diag(10,d,d)
n0 = 10
initmu = matrix(0,d,k)
initsigma = array(n0*V0,dim = c(d,d,k))

priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
priors$beta <- c(0,2*betacritic)

mh <- list(bandwidth=0.1)
```

```{r}
iter = 10000
burnin = 5000
results <- MCMCPotts(data,neigh,block,iter,burnin,priors,mh)
```

```{r}
clustering <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering[pixels[i,1],pixels[i,2]] = which.max(results$alloc[i,])
}
print(table(clustering))
```

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering, border=NA,asp = TRUE,col= turbo(k),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```

```{r}
plot(pcascore1vec,pcascore2vec,col = clustering[pixels[,1],pixels[,2]])
```

```{r}
mu1chain = mcmc(t(results$mu[,1,]))
mu2chain = mcmc(t(results$mu[,2,]))
mu3chain = mcmc(t(results$mu[,3,]))
```

```{r}
par(mfrow = c(d,d))
plot(mu1chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu1chain,auto.layout = FALSE)
plot(mu2chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu2chain,auto.layout = FALSE)
plot(mu3chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu3chain,auto.layout = FALSE)
```


```{r}
trace1 = rep(0,iter)
trace2 = rep(0,iter)
trace3 = rep(0,iter)

for( i in 1:iter){
  trace1[i] = sum(diag(results$sigmas[i][[1]][,,1]))
  trace2[i] = sum(diag(results$sigmas[i][[1]][,,2]))
  trace3[i] = sum(diag(results$sigmas[i][[1]][,,3]))
}

tr1chain = mcmc(trace1)
tr2chain = mcmc(trace2)
tr3chain = mcmc(trace3)

par(mfrow = c(3,2))
plot(tr1chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr1chain,auto.layout = FALSE)
plot(tr2chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr2chain,auto.layout = FALSE)
plot(tr3chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr3chain,auto.layout = FALSE)
```

```{r}
betachain = mcmc(results$beta)
plot(betachain)
autocorr.plot(betachain)
```

## K = 6

we start with non informative priors for the parameters

```{r}
k = 6
betacritic = log(1 + sqrt(k))
d = 2
B0 = diag(100,d,d)
V0 = diag(10,d,d)
n0 = 10
initmu = matrix(0,d,k)
initsigma = array(n0*V0,dim = c(d,d,k))

priors <- list()
priors$k <- k
priors$mu <- matrix(0,d,k)
priors$mu.sigma <- array(B0,dim = c(d,d,k))
priors$sigma.V0 <- array(V0,dim = c(d,d,k))
priors$sigma.n0 <- rep(n0,k)
priors$beta <- c(0,2*betacritic)

mh <- list(bandwidth=0.1)
```


```{r}
iter = 10000
burnin = 5000
results <- MCMCPotts(data,neigh,block,iter,burnin,priors,mh)
```

```{r}
clustering <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering[pixels[i,1],pixels[i,2]] = which.max(results$alloc[i,])
}
print(table(clustering))
```

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering, border=NA,asp = TRUE,col= turbo(k),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```

```{r}
plot(pcascore1vec,pcascore2vec,col = clustering[pixels[,1],pixels[,2]])
```

```{r}
mu1chain = mcmc(t(results$mu[,1,]))
mu2chain = mcmc(t(results$mu[,2,]))
mu3chain = mcmc(t(results$mu[,3,]))
mu4chain = mcmc(t(results$mu[,4,]))
mu5chain = mcmc(t(results$mu[,5,]))
mu6chain = mcmc(t(results$mu[,6,]))
```


```{r}
par(mfrow = c(d,d))
plot(mu1chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu1chain,auto.layout = FALSE)
plot(mu2chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu2chain,auto.layout = FALSE)
plot(mu3chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu3chain,auto.layout = FALSE)
plot(mu4chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu4chain,auto.layout = FALSE)
plot(mu5chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu5chain,auto.layout = FALSE)
plot(mu6chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(mu6chain,auto.layout = FALSE)
```

```{r}
trace1 = rep(0,iter)
trace2 = rep(0,iter)
trace3 = rep(0,iter)
trace4 = rep(0,iter)
trace5 = rep(0,iter)
trace6 = rep(0,iter)

for( i in 1:iter){
  trace1[i] = sum(diag(results$sigmas[i][[1]][,,1]))
  trace2[i] = sum(diag(results$sigmas[i][[1]][,,2]))
  trace3[i] = sum(diag(results$sigmas[i][[1]][,,3]))
  trace4[i] = sum(diag(results$sigmas[i][[1]][,,4]))
  trace5[i] = sum(diag(results$sigmas[i][[1]][,,5]))
  trace6[i] = sum(diag(results$sigmas[i][[1]][,,6]))
}

tr1chain = mcmc(trace1)
tr2chain = mcmc(trace2)
tr3chain = mcmc(trace3)
tr4chain = mcmc(trace4)
tr5chain = mcmc(trace5)
tr6chain = mcmc(trace6)

par(mfrow = c(2,2))
plot(tr1chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr1chain,auto.layout = FALSE)
plot(tr2chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr2chain,auto.layout = FALSE)
plot(tr3chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr3chain,auto.layout = FALSE)
plot(tr4chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr4chain,auto.layout = FALSE)
plot(tr5chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr5chain,auto.layout = FALSE)
plot(tr6chain,density = FALSE,auto.layout = FALSE)
autocorr.plot(tr6chain,auto.layout = FALSE)
```

```{r}
betachain = mcmc(results$beta)
plot(betachain)
autocorr.plot(betachain)
```
