---title: "Potts model Peptidi, beta fisso"
author: Simone Colombara, Alessia Cotroneo, Francesco De Caro, Riccardo Morandi, Chiara   Schembri,
  Alfredo Zapiola
date: "2022-12-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rayshader)
library(patchwork)
library(ggplot2)
library(plot.matrix)
library(bayesImageS)
library(coda)
library(viridis)
```


```{r}
D = read.table("C:/Users/Francesco/OneDrive - Politecnico di Milano/Desktop/154 variabili/101_peptidi-PreProcessed-IM-Step1-Step2-Step4-Step5-101.txt")
D0 = D
D0[is.na(D0)] = 0

pixels = read.table("C:/Users/Francesco/OneDrive - Politecnico di Milano/Desktop/154 variabili/101_peptidi-PreProcessed-XYCoordinates-Step1-Step2-Step4-Step5-101.txt")
colnames(D0) = substr(colnames(D0),1,4)
colnames(pixels) = c("x","y")

Data_long = as_tibble(data.frame( pixels, D0 ))
max_number_of_pixels = apply(Data_long[,1:2],2,max)

pca = princomp(D0)
pcascore1 <-as.vector(pca$scores[,1])
```

```{r}
pcascore1vec = pca$scores[,1]
max(pcascore1vec)
min(pcascore1vec)
mean(pcascore1vec)
sd(pcascore1vec)
library(viridis)
```

```{r}
par(mfrow=c(1,2))
hist(pcascore1vec,main = "first pc score",breaks = 100)
plot(density(pcascore1vec),main = "kernel density first pc score")
```



```{r}
mask <-matrix(0,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  mask[pixels[i,1],pixels[i,2]] = 1
}

neigh <- getNeighbors(mask = mask, c(2,2,0,0))
block <- getBlocks(mask = mask, 2)

```

##GIBBSPOTTS
#k=3


```{r, results ='hide'}
q <- 3
betacritic = log(1 + sqrt(q))

mu <- rep(0,q)
sd <- rep(2,q)

priors <- list()
priors$k <- q
priors$mu <- rep(0,q)
priors$mu.sd <- rep(70,q)
priors$sigma <- rep(4,q)
priors$sigma.nu <- rep(0.1,q)

prioraux <- priors
prioraux$beta <- c(0,1)
mh<- list(algorithm="pseudolikelihood", bandwidth=1)
resaux<- mcmcPotts(pcascore1, neigh, block, prioraux, mh, 1, 0)

labels <- resaux$z # this is just to get the correct data type 
#random initialization of the labels
for(i in 1:dim(pixels)[1]){
  labels[i,] = rep(0,q)
  labels[i,which.max(runif(q))] = 1
}
```


```{r, results='hide'}
resbc <- gibbsPotts(pcascore1, labels, betacritic, mu, sd, neigh, block, priors, niter = 10000)
```

```{r, results='hide'}
clusteringgibbs <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clusteringgibbs[pixels[i,1],pixels[i,2]] = which.max(resbc$z[i,])
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clusteringgibbs, border=NA,asp = TRUE,col = viridis(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```
#no constraint on betacritic
```{r, results='hide'}
resbc <- gibbsPotts(pcascore1, labels, 2*betacritic, mu, sd, neigh, block, priors, niter = 10000)
```


```{r, results='hide'}
clusteringgibbs <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clusteringgibbs[pixels[i,1],pixels[i,2]] = which.max(resbc$z[i,])
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clusteringgibbs, border=NA,asp = TRUE,col = turbo(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```

#k=4

```{r, results ='hide'}
q <- 4
betacritic = log(1 + sqrt(q))

mu <- rep(0,q)
sd <- c(rep(10,q-1),0.1)

priors <- list()
priors$k <- q
priors$mu <- c(rep(0,q-1),100)
priors$mu.sd <- c(rep(20,q-1),0.1)
priors$sigma <- c(rep(4,q-1),0.1)
priors$sigma.nu <- rep(0.1,q)

prioraux <- priors
prioraux$beta <- c(0,1)
mh<- list(algorithm="pseudolikelihood", bandwidth=1)
resaux<- mcmcPotts(pcascore1, neigh, block, prioraux, mh, 1, 0)

labels <- resaux$z # this is just to get the correct data type 
#random initialization of the labels
for(i in 1:dim(pixels)[1]){
  labels[i,] = rep(0,q)
  labels[i,which.max(runif(q))] = 1
}
```


```{r, results='hide'}
res4 <- gibbsPotts(pcascore1, labels, betacritic, mu, sd, neigh, block, priors, niter = 10000)
```


```{r, results='hide'}
clusteringgibbs4 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clusteringgibbs4[pixels[i,1],pixels[i,2]] = which.max(res4$z[i,])
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clusteringgibbs4, border=NA,asp = TRUE,col = turbo(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```

#no constraint on betacritic
```{r, results='hide'}
resbc <- gibbsPotts(pcascore1, labels, 2*betacritic, mu, sd, neigh, block, priors, niter = 10000)
```


```{r, results='hide'}
clusteringgibbs4 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clusteringgibbs4[pixels[i,1],pixels[i,2]] = which.max(resbc$z[i,])
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clusteringgibbs4, border=NA,asp = TRUE,col = turbo(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```

##PSEUDOLIKELIHOOD

```{r}
mask <-matrix(0,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  mask[pixels[i,1],pixels[i,2]] = 1
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(mask, border=NA,asp = TRUE)
```


# K = 3

```{r}
q <- 3
betacritic = log(1 + sqrt(q))
neigh <- getNeighbors(mask = mask, c(2,2,0,0))
block <- getBlocks(mask = mask, 2)

priors <- list()
priors$k <- q
priors$mu <- c(-15,0,40)
priors$mu.sd <- rep(2,q)
priors$sigma <- rep(2,q)
priors$sigma.nu <- rep(0.5,q)
priors$beta <- c(0,betacritic)

mh <- list(algorithm="pseudolikelihood", bandwidth=1)
```


```{r, results='hide'}
res <- mcmcPotts(pcascore1, neigh, block, priors, mh, 10000, 5000)
```

## chain analysis 

```{r}
muchain = mcmc(res$mu)
varnames(muchain)<-c("mu_1","mu_2","mu_3")
sigmachain = mcmc(res$sigma)
varnames(sigmachain)<-c("sigma_1","sigma_2","sigma_3")
betachain = mcmc(res$beta)
varnames(betachain)<-c("beta")
sumchain  = mcmc(res$sum)
varnames(sumchain)<-c("sum")
```

### mu

```{r}

summary(muchain)
batchSE(muchain)
effectiveSize(muchain)
rejectionRate(muchain)
# still need to understand this since i am using adaptive mh alg
plot(muchain)

par(mfrow=c(1,3))
autocorr.plot(muchain,auto.layout=FALSE)

crosscorr.plot(muchain)
```

### sigma 

```{r}

summary(sigmachain)
batchSE(sigmachain)
effectiveSize(sigmachain)

plot(sigmachain)

par(mfrow=c(1,3))
autocorr.plot(sigmachain,auto.layout=FALSE)
```

### sum

```{r}
summary(sumchain)
batchSE(sumchain)
effectiveSize(sumchain)

plot(sumchain)

autocorr.plot(sumchain)
```

### beta

```{r}
summary(betachain)
batchSE(betachain)
effectiveSize(betachain)

plot(betachain)

autocorr.plot(betachain)
```

## plot

```{r}
clustering <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering[pixels[i,1],pixels[i,2]] = res$e[i]
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering, border=NA,asp = TRUE,col = rainbow)
```


## I want to see what happens if i remove the constraints on beta 
```{r, results='hide'}
priors2 <- list()
priors2$k <- q
priors2$mu <- c(-15,0,40)
priors2$mu.sd <- rep(2,q)
priors2$sigma <- rep(2,q)
priors2$sigma.nu <- rep(0.5,q)
priors2$beta <- c(0,2*betacritic)

mh2 <- list(algorithm="pseudolikelihood", bandwidth=1)

res2 <- mcmcPotts(pcascore1, neigh, block, priors2, mh2, 10000, 5000)

clustering2 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering2[pixels[i,1],pixels[i,2]] = res2$e[i]
}
```

## chain analysis

```{r}

muchain2 = mcmc(res2$mu)
varnames(muchain2)<-c("mu_1","mu_2","mu_3")
sigmachain2 = mcmc(res2$sigma)
varnames(sigmachain2)<-c("sigma_1","sigma_2","sigma_3")
betachain2 = mcmc(res2$beta)
varnames(betachain2)<-c("beta")
sumchain2  = mcmc(res2$sum)
varnames(sumchain2)<-c("sum")

```

### mu

```{r}

summary(muchain2)
batchSE(muchain2)
effectiveSize(muchain2)
rejectionRate(muchain2)
# still need to understand this since i am using adaptive mh alg

plot(muchain2)

par(mfrow=c(1,3))
autocorr.plot(muchain2,auto.layout=FALSE)

crosscorr.plot(muchain2)

```

### sigma

```{r}
summary(sigmachain2)
batchSE(sigmachain2)
effectiveSize(sigmachain2)

plot(sigmachain2)

par(mfrow=c(1,3))
autocorr.plot(sigmachain2,auto.layout=FALSE)
```

###sum

```{r}
summary(sumchain2)
batchSE(sumchain2)
effectiveSize(sumchain2)

plot(sumchain2)

autocorr.plot(sumchain2)
```

###beta

```{r}
summary(betachain2)
batchSE(betachain2)
effectiveSize(betachain2)

plot(betachain2)

autocorr.plot(betachain2)

```


```{r}
print(table(clustering2))
```

## plot

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering2, border=NA,asp = TRUE,col = rainbow)
```

# K = 4

```{r}
q <- 4
betacritic = log(1 + sqrt(q))
neigh <- getNeighbors(mask = mask, c(2,2,0,0))
block <- getBlocks(mask = mask, 2)

priors <- list()
priors$k <- q
priors$mu <- c(-15,0,20,40)
priors$mu.sd <- rep(2,q)
priors$sigma <- rep(2,q)
priors$sigma.nu <- rep(0.5,q)
priors$beta <- c(0,betacritic)

mh <- list(algorithm="pseudolikelihood", bandwidth=1)
```


```{r, results='hide'}
res4 <- mcmcPotts(pcascore1, neigh, block, priors, mh, 10000, 5000)
```

## chain analysis 

```{r}
muchain = mcmc(res4$mu)
varnames(muchain)<-c("mu_1","mu_2","mu_3", "mu_4")
sigmachain = mcmc(res4$sigma)
varnames(sigmachain)<-c("sigma_1","sigma_2","sigma_3", "sima_4")
betachain = mcmc(res4$beta)
varnames(betachain)<-c("beta")
sumchain  = mcmc(res4$sum)
varnames(sumchain)<-c("sum")
```

### mu

```{r}

summary(muchain)
batchSE(muchain)
effectiveSize(muchain)
rejectionRate(muchain)
# still need to understand this since i am using adaptive mh alg
plot(muchain)

par(mfrow=c(1,3))
autocorr.plot(muchain,auto.layout=FALSE)

crosscorr.plot(muchain)
```

### sigma 

```{r}

summary(sigmachain)
batchSE(sigmachain)
effectiveSize(sigmachain)

plot(sigmachain)

par(mfrow=c(1,3))
autocorr.plot(sigmachain,auto.layout=FALSE)
```

### sum

```{r}
summary(sumchain)
batchSE(sumchain)
effectiveSize(sumchain)

plot(sumchain)

autocorr.plot(sumchain)
```

### beta

```{r}
summary(betachain)
batchSE(betachain)
effectiveSize(betachain)

plot(betachain)

autocorr.plot(betachain)
```

## plot

```{r}
clustering4 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering4[pixels[i,1],pixels[i,2]] = res4$e[i]
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering4, border=NA,asp = TRUE,col = rainbow)
```


## I want to see what happens if i remove the constraints on beta 
```{r, results='hide'}
priors2 <- list()
priors2$k <- q
priors2$mu <- c(-15,0,20,40)
priors2$mu.sd <- rep(2,q)
priors2$sigma <- rep(2,q)
priors2$sigma.nu <- rep(0.5,q)
priors2$beta <- c(0,2*betacritic)

mh2 <- list(algorithm="pseudolikelihood", bandwidth=1)

res2 <- mcmcPotts(pcascore1, neigh, block, priors2, mh2, 10000, 5000)

clustering2 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering2[pixels[i,1],pixels[i,2]] = res2$e[i]
}
```

## chain analysis

```{r}

muchain2 = mcmc(res2$mu)
varnames(muchain2)<-c("mu_1","mu_2","mu_3","mu_4")
sigmachain2 = mcmc(res2$sigma)
varnames(sigmachain2)<-c("sigma_1","sigma_2","sigma_3", "sigma_4")
betachain2 = mcmc(res2$beta)
varnames(betachain2)<-c("beta")
sumchain2  = mcmc(res2$sum)
varnames(sumchain2)<-c("sum")

```

### mu

```{r}

summary(muchain2)
batchSE(muchain2)
effectiveSize(muchain2)
rejectionRate(muchain2)
# still need to understand this since i am using adaptive mh alg
par(mfrow=c(2,2))
plot(muchain2, auto.layout = FALSE)

par(mfrow=c(2,2))
autocorr.plot(muchain2,auto.layout=FALSE)


```

### sigma

```{r}
summary(sigmachain2)
batchSE(sigmachain2)
effectiveSize(sigmachain2)
par(mfrow=c(2,2))
plot(sigmachain2,auto.layout=FALSE)

par(mfrow=c(2,2))
autocorr.plot(sigmachain2,auto.layout=FALSE)
```

###sum

```{r}
summary(sumchain2)
batchSE(sumchain2)
effectiveSize(sumchain2)

plot(sumchain2)

autocorr.plot(sumchain2)
```

###beta

```{r}
summary(betachain2)
batchSE(betachain2)
effectiveSize(betachain2)

plot(betachain2)

autocorr.plot(betachain2)

```


```{r}
print(table(clustering2))
```

## plot

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering2, border=NA,asp = TRUE,col = rainbow)
```
