---
title: "Data_Inspection"
author: "Riccardo Morandi"
date: "2022-10-19"
output: html_document
---
```{r libraries}
knitr::opts_chunk$set(fig.width = 8,fig.height = 6)
library(tidyverse)
library(rayshader)
library(patchwork)
library(skimr)
library(visdat)
library(ggplot2)
library(GGally)
library(corrplot)
```

# A first inspection of the dataset with 78 variables

## importing the data




```{r, read data}
D = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-IM-Step1-Step2-Step4-Step5-101.txt")
sum(is.na(D))
```
 the numbers of na is substantial
```{r }
 vis_miss(D,warn_large_data = FALSE)
```
the missing data is about 45% 
 
```{r}
skim(D)
```

we observe that the missing data is not uniform in the mz, there are some values for which only 20 - 30% of the pixel have a value, and this tends to be small

we replace the missing data with 0 since it means the data for that mz was under threshold

```{r}
D0 = D
D0[is.na(D0)] = 0
```

# preliminary plotts
look at the distribution and correlation of the data, beginning with similar mz
probably ggpairs is  not the most efficient, did not change the labels 


```{r, warning = FALSE}
ggpairs(D,columns = 1:7)

ggpairs(D,columns = 8:14)

ggpairs(D,columns = 15:21)

ggpairs(D,columns = 16:28)

ggpairs(D,columns = 28:35)

ggpairs(D,columns = 36:42)

ggpairs(D,columns = 43:49)

ggpairs(D,columns = 50:56)
```

we see a lot of correlation in the data especially below


```{r, warning = FALSE}
ggpairs(D,columns = 57:63)

ggpairs(D,columns = 64:70)

ggpairs(D,columns = 70:78)
```

* do the zeros added contribute a lot to the correlation?

* the peaks that we have around 800 mz are all correlated and are basically the same information
* this high correlation explains the performance of pca

# correlation matrix
```{r}
cm <- cor(D0)
colnames(D0) = substr(colnames(D0),1,4)
corrplot(cm, method = "color")
```
the blue blob is the observation from above

```{r}
pixels = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-XYCoordinates-Step1-Step2-Step4-Step5-101.txt")
colnames(D0) = substr(colnames(D0),1,4)
colnames(pixels) = c("x","y")
```

Create the datasets we will need:

- `Data_long`: contains the 18k pixels on the rows. The first two columns are the coordinates, the remaining 78 are the values recorded for different `m.z`.

- `Data_array`: contains a cube with the 78 slices on the third dimension. Not all the pixels are recored in each rectangle: beware of `NA`'s.



```{r}
Data_long            = as_tibble(data.frame( pixels, D0 ))
max_number_of_pixels = apply(Data_long[,1:2],2,max)

Data_array = matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])

Data_array = array(NA,c(max_number_of_pixels[1],max_number_of_pixels[2],ncol(D0)))

sum(is.na(D0))

# there must be a better way to do this, but it's sunday morning, please be patient...
for(k in 1:ncol(D0)){
  for(i in 1:nrow(Data_long)){
  Data_array[Data_long$x[i],Data_long$y[i],k] = D0[i,k]
  }
}

dim(Data_array)

Data_very_long = reshape2::melt(Data_long,c("x","y")) %>% mutate(pixel_ind = paste0(x,"_",y), value_ind = rep(1:nrow(Data_long),ncol(D0)))
Data_very_long = Data_very_long %>% group_by(pixel_ind) %>% mutate(n = row_number()) %>% ungroup() %>% mutate(mz = as.numeric(substr(variable,2,4)))

```


```{r}
Data_very_long = reshape2::melt(Data_long,c("x","y")) %>% mutate(pixel_ind = paste0(x,"_",y), value_ind = rep(1:nrow(Data_long),ncol(D0)))

Data_very_long = Data_very_long %>% group_by(pixel_ind) %>% mutate(n = row_number()) %>% ungroup() %>% mutate(mz = as.numeric(substr(variable,2,4)))


# subsampling to get a faster plot and not drain memory
sub_ind = sample(unique(Data_very_long$pixel_ind),100)
# just to get the gist:
ggplot(Data_very_long %>% filter(pixel_ind %in% sub_ind))+
  geom_path(aes(x = mz, y = value, 
                col=pixel_ind, 
                group = pixel_ind),alpha=.5)+theme_bw()+theme(legend.position = "none")+xlab("m.z")+scale_color_viridis_d(option = "A")+
  scale_x_continuous(n.breaks = 20)
```

# investigating the different peaks

## the first mz don't contain a lot of info, quite noisy


```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X437))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X442))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X447))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X448))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X449))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X465))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```


# we investigate the spike arround 500

```{r,echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X497))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X506))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X511))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X513))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X524))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X525))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6
```

the spike is in mz 511 on the edge  
- possible problem of the instrument on the edge of the brain 
- outlier ?
the rest is just noise

# spike arround 775

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X775))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X775))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X778))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X779))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X780))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X793))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6
```


the spike is relative to 778 779 780

# spike arround 800 and 825

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X806))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X807))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X808))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X822))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X823))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X836))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6
```

very similar to each other

# after 850

```{r, echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X862))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X863))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X888))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X889))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X890))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X904))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6
```

this are all the same, it is the high correlated blob in the correlation matrix

# there is another peak around 900
```{r,echo = FALSE}
P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X904))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X906))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X907))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X908))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X933))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X934))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6
```

the peak is in 906 907 908 which are very similar

the rest is not that itresting, very low values resamble kind on veramble the previuus structure but basically noise

# A comprehensive look at the peaks

```{r, echo = FALSE}

P1 = ggplot(Data_long)+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = X511))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P2 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X778))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P3 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X807))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P4 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X888))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P5 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X890))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
P6 = ggplot(Data_long)+
  geom_tile(aes(x=x,y=y,fill = X906))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

P1+P2+P3+P4+P5+P6

```


# exept for the 511 which is in the edge, the peaks are all correlarend and therefore are all the same information

# PCA

```{r}
pca = princomp(D0)
plot(pca)
summary(pca)
```
the pca works well because of the correlation

```{r}


PCA1 = ggplot(Data_long %>% mutate(pca1 = pca$scores[,1]))+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = pca1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")
PCA2 = ggplot(Data_long %>% mutate(pca2 = pca$scores[,2]))+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = pca2))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

PCA1+PCA2
```


# Note: we are always using the same info 


```{r, echo = FALSE}
PCA1 + P6
```

if we invert the colormap

```{r}
PCA1v2 = ggplot(Data_long %>% mutate(pca1 = pca$scores[,1]))+ theme_bw()+
  geom_tile(aes(x=x,y=y,fill = -1*pca1))+scale_fill_viridis_c(option = "A",na.value = "red")+
  theme_void()+theme(legend.position = "bottom")

PCA1v2 + P6
```

the data is clearly dominated by this peaks that are highly correlated  and spatially correlated as well

# Plotting the principal components

first component

```{r,echo = FALSE}
b1 = data.frame(pca$loadings[,1])
b1$mz = unique(Data_very_long$mz)
ggplot(b1)+ geom_path(aes(x = mz, y = -1*pca.loadings...1.))+theme_bw()+theme(legend.position = "none")+xlab("m.z")+scale_color_viridis_d(option = "A")+
  scale_x_continuous(n.breaks = 20)
```
second component

```{r,echo = FALSE}
b2 = data.frame(pca$loadings[,2])
b2$mz = unique(Data_very_long$mz)
ggplot(b2)+ geom_path(aes(x = mz, y = pca.loadings...2.))+theme_bw()+theme(legend.position = "none")+xlab("m.z")+scale_color_viridis_d(option = "A")+
  scale_x_continuous(n.breaks = 20)

```


# do we need to be smarter and explore the non dominant factors or is enough? 

```{r}
sessionInfo()
```

