---
title: "Beta Identification"
author: Simone Colombara, Alessia Cotroneo, Francesco De Caro, Riccardo Morandi, Chiara   Schembri,
  Alfredo Zapiola
date: "2022-12-13"
output: html_document
---
```{r libraries, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rayshader)
library(patchwork)
library(ggplot2)
library(plot.matrix)
library(bayesImageS)
library(coda)
library(viridis)
```

importing data pixel structure

```{r}
D = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-IM-Step1-Step2-Step4-Step5-101.txt")
D0 = D
D0[is.na(D0)] = 0

pixels = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-XYCoordinates-Step1-Step2-Step4-Step5-101.txt")
colnames(D0) = substr(colnames(D0),1,4)
colnames(pixels) = c("x","y")

Data_long = as_tibble(data.frame( pixels, D0 ))
max_number_of_pixels = apply(Data_long[,1:2],2,max)

```

building auxiliary mask to pass just the measured pixels to the functions

```{r}
mask <-matrix(0,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  mask[pixels[i,1],pixels[i,2]] = 1
}

```

# beta = betacritic

## simulating data

we first decide to generate synthetic data using the Swendsen & Wang algorithm, starting with a full matrix to get reliable patterns

```{r}
full_mask <-matrix(1,max_number_of_pixels[2],max_number_of_pixels[2])

full_neigh <- getNeighbors(mask = full_mask, c(2,2,0,0))
full_blocks <- getBlocks(mask = full_mask, 2)

```

```{r}
k = 3
betacritic = log(1 + sqrt(k))

sw_res <- swNoData(betacritic, k, full_neigh, full_blocks, niter = 1000, random = TRUE)

sinth_clust <- matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:max_number_of_pixels[1]){
  for(j in 1:max_number_of_pixels[2]){
  sinth_clust[i,j] = which.max(sw_res$z[(i-1)*max_number_of_pixels[2]+j,])
  }
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(sinth_clust, border=NA,asp = TRUE,col = viridis,axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)

```

```{r}
allocs = rep(0,dim(pixels)[1])
for(i in 1:dim(pixels)[1]){
  allocs[i] = sinth_clust[pixels[i,1],pixels[i,2]] 
}

plot_allocs =  matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  plot_allocs[pixels[i,1],pixels[i,2]] = allocs[i]  
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(plot_allocs, border=NA,asp = TRUE,col = viridis,axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```

generating the data 

```{r}
mu1 = -4
mu2 = 0
mu3 = 5

s1 = 3
s2 = 2
s3 = 1

t = table(allocs)

n1 = t[1]
n2 = t[2]
n3 = t[3]

x1 = rnorm(n1,mean = mu1,sd = s1)
x2 = rnorm(n2,mean = mu2,sd = s2)
x3 = rnorm(n3,mean = mu3,sd = s3)

c1 = 1
c2 = 1
c3 = 1

data = rep(0,length(allocs))
for(i in 1:length(data)){
  if(allocs[i]==1){
      data[i] = x1[c1]
      c1 = c1+1
    }
    else if(allocs[i]==2){
      data[i] = x2[c2]
      c2 = c2+1
    }
    else if (allocs[i]==3){
      data[i] = x3[c3]
      c3 = c3+1
    }
}

plot_data =  matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  plot_data[pixels[i,1],pixels[i,2]] = data[i]  
}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(plot_data, border=NA,asp = TRUE,col = viridis,axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```


```{r}
q <- 3
betacritic = log(1 + sqrt(q))
neigh <- getNeighbors(mask = mask, c(2,2,0,0))
block <- getBlocks(mask = mask, 2)

priors <- list()
priors$k <- q
priors$mu <- c(-5,0,5)
priors$mu.sd <- rep(3,q)
priors$sigma <- rep(2,q)
priors$sigma.nu <- rep(1.5,q)
priors$beta <- c(0,2*betacritic)

mh <- list(algorithm="pseudolikelihood", bandwidth=1)
```


```{r, results='hide'}
res <- mcmcPotts(data, neigh, block, priors, mh, 10000, 5000)
```

```{r}
muchain = mcmc(res$mu)
varnames(muchain)<-c("mu_1","mu_2","mu_3")
sigmachain = mcmc(res$sigma)
varnames(sigmachain)<-c("sigma_1","sigma_2","sigma_3")
betachain = mcmc(res$beta)
varnames(betachain)<-c("beta")
sumchain  = mcmc(res$sum)
varnames(sumchain)<-c("sum")
```

### beta

```{r}
summary(betachain)
batchSE(betachain)
effectiveSize(betachain)

plot(betachain)

autocorr.plot(betachain)
```
we can correctly identify the beta eve if beta = betacritic this suggests that the beta in our data is above betacritic

## plot

```{r}
clustering <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering[pixels[i,1],pixels[i,2]] = res$e[i]
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering, border=NA,asp = TRUE,col = rainbow)
```

```{r}
diff = 0

for(i in 1:dim(pixels)[1]){
  if(clustering[pixels[i,1],pixels[i,2]] != plot_allocs[pixels[i,1],pixels[i,2]])
    diff = diff+1
}
diff 
error = diff/dim(pixels)[1]
error

```

# beta > betacritic

## simulating data

we first decide to generate synthetic data using the Swendsen & Wang algorithm, starting with a full matrix to get reliable patterns

```{r}
k = 3
betacritic = log(1 + sqrt(k))
beta = 1.001*betacritic
sw_res <- swNoData(beta, k, full_neigh, full_blocks, niter = 1000, random = TRUE)

sinth_clust <- matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:max_number_of_pixels[1]){
  for(j in 1:max_number_of_pixels[2]){
  sinth_clust[i,j] = which.max(sw_res$z[(i-1)*max_number_of_pixels[2]+j,])
  }
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(sinth_clust, border=NA,asp = TRUE,col = viridis,axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)

```

```{r}
allocs = rep(0,dim(pixels)[1])
for(i in 1:dim(pixels)[1]){
  allocs[i] = sinth_clust[pixels[i,1],pixels[i,2]] 
}

plot_allocs =  matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  plot_allocs[pixels[i,1],pixels[i,2]] = allocs[i]  
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(plot_allocs, border=NA,asp = TRUE,col = viridis,axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```

generating the data 
even a little step witth beta creates a lot of uniformity in the data

```{r}
mu1 = -4
mu2 = 0
mu3 = 5

s1 = 3
s2 = 2
s3 = 1

t = table(allocs)

n1 = t[1]
n2 = t[2]
n3 = t[3]

x1 = rnorm(n1,mean = mu1,sd = s1)
x2 = rnorm(n2,mean = mu2,sd = s2)
x3 = rnorm(n3,mean = mu3,sd = s3)

c1 = 1
c2 = 1
c3 = 1

data = rep(0,length(allocs))
for(i in 1:length(data)){
  if(allocs[i]==1){
      data[i] = x1[c1]
      c1 = c1+1
    }
    else if(allocs[i]==2){
      data[i] = x2[c2]
      c2 = c2+1
    }
    else if (allocs[i]==3){
      data[i] = x3[c3]
      c3 = c3+1
    }
}

plot_data =  matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  plot_data[pixels[i,1],pixels[i,2]] = data[i]  
}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(plot_data, border=NA,asp = TRUE,col = viridis,axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```


```{r}
q <- 3
betacritic = log(1 + sqrt(q))
neigh <- getNeighbors(mask = mask, c(2,2,0,0))
block <- getBlocks(mask = mask, 2)

priors <- list()
priors$k <- q
priors$mu <- c(-5,0,5)
priors$mu.sd <- rep(3,q)
priors$sigma <- rep(2,q)
priors$sigma.nu <- rep(1.5,q)
priors$beta <- c(0,2*betacritic)

mh <- list(algorithm="pseudolikelihood", bandwidth=1)
```


```{r, results='hide'}
res <- mcmcPotts(data, neigh, block, priors, mh, 10000, 5000)
```

```{r}
muchain = mcmc(res$mu)
varnames(muchain)<-c("mu_1","mu_2","mu_3")
sigmachain = mcmc(res$sigma)
varnames(sigmachain)<-c("sigma_1","sigma_2","sigma_3")
betachain = mcmc(res$beta)
varnames(betachain)<-c("beta")
sumchain  = mcmc(res$sum)
varnames(sumchain)<-c("sum")
```

### beta

```{r}
summary(betachain)
batchSE(betachain)
effectiveSize(betachain)

plot(betachain)

autocorr.plot(betachain)
```
we can correctly identify the beta eve if beta = betacritic this suggests that the beta in our data is above betacritic

## plot

```{r}
clustering <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering[pixels[i,1],pixels[i,2]] = which.max(res$alloc[i,])
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering, border=NA,asp = TRUE,col = rainbow)
```

we can still obtain the right pattern

```{r}
diff = 0

for(i in 1:dim(pixels)[1]){
  if(clustering[pixels[i,1],pixels[i,2]] != plot_allocs[pixels[i,1],pixels[i,2]])
    diff = diff+1
}
diff 
error = diff/dim(pixels)[1]
error

```