---
title: "Beta investigation"
author: Simone Colombara, Alessia Cotroneo, Francesco De Caro, Riccardo Morandi, Chiara Schembri,
  Alfredo Zapiola
date: "2022-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lattice)
library(tidyverse)
library(rayshader)
library(patchwork)
library(bayesImageS)
```



```{r}
pixels = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-XYCoordinates-Step1-Step2-Step4-Step5-101.txt")
colnames(pixels) = c("x","y")
D = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-IM-Step1-Step2-Step4-Step5-101.txt")
D0 = D
D0[is.na(D0)] = 0
Data_long = as_tibble(data.frame( pixels, D0 ))
max_number_of_pixels = apply(Data_long[,1:2],2,max)
```

we simulate data from different geometries in order to asses whether the beta critic is affected by it

# square geomerty

```{r }
mask <-matrix(1,max_number_of_pixels[2],max_number_of_pixels[2])

neigh <- getNeighbors(mask = mask, c(2,2,0,0))
block <- getBlocks(mask = mask, 2)

```


```{r }
beta <- seq(0, 2, by = 0.05)
iter <- 1000
burn <- 800
n <- sum(mask)
k <- 3
bcrit <- log(1 + sqrt(k))
samp.bIS <- matrix(nrow = length(beta), ncol = iter - burn)
```


```{r }
for (i in 1:length(beta)) {
  result <- swNoData(beta[i],k,neigh,block,iter)
  samp.bIS[i,] <- result$sum[(burn+1):iter]
}
```


```{r }
s_z <- samp.bIS
s_x <- rep(beta,times=iter-burn)
sd_s <- rep(0,dim(s_z)[1])
for(i in 1:dim(s_z)[1]){
  sd_s[i] = sd(s_z[i,])
}

par(mfrow=c(1,2))
plot(s_x,s_z,xlab=expression(beta),ylab=expression(S(z)))
abline(v=bcrit,col="red")

plot(beta,sd_s,xlab=expression(beta),ylab=expression(sigma[S(z)]))
abline(v=bcrit,col="red")

```

# rectagular geometry

```{r }
mask <-matrix(1,max_number_of_pixels[1],max_number_of_pixels[2])

neigh <- getNeighbors(mask = mask, c(2,2,0,0))
block <- getBlocks(mask = mask, 2)

```


```{r }
beta <- seq(0, 2, by = 0.05)
iter <- 1000
burn <- 800
n <- sum(mask)
k <- 3
bcrit <- log(1 + sqrt(k))
sampr.bIS <- matrix(nrow = length(beta), ncol = iter - burn)
```


```{r }
for (i in 1:length(beta)) {
  result <- swNoData(beta[i],k,neigh,block,iter)
  samp.bIS[i,] <- result$sum[(burn+1):iter]
}
```


```{r }
s_zr <- samp.bIS
s_xr <- rep(beta,times=iter-burn)
sd_sr <- rep(0,dim(s_zr)[1])
for(i in 1:dim(s_zr)[1]){
  sd_sr[i] = sd(s_zr[i,])
}

par(mfrow=c(1,2))
plot(s_xr,s_zr,xlab=expression(beta),ylab=expression(S(z)))
abline(v=bcrit,col="red")

plot(beta,sd_sr,xlab=expression(beta),ylab=expression(sigma[S(z)]))
abline(v=bcrit,col="red")

```


# brain geometry

```{r }
mask <-matrix(0,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  mask[pixels[i,1],pixels[i,2]] = 1
}

neigh <- getNeighbors(mask = mask, c(2,2,0,0))
block <- getBlocks(mask = mask, 2)

```


```{r }
beta <- seq(0, 2, by = 0.05)
iter <- 1000
burn <- 800
n <- sum(mask)
k <- 3
bcrit <- log(1 + sqrt(k))
sampb.bIS <- matrix(nrow = length(beta), ncol = iter - burn)
```


```{r }
for (i in 1:length(beta)) {
  result <- swNoData(beta[i],k,neigh,block,iter)
  samp.bIS[i,] <- result$sum[(burn+1):iter]
}
```


```{r }
s_zb <- samp.bIS
s_xb <- rep(beta,times=iter-burn)
sd_sb <- rep(0,dim(s_z)[1])
for(i in 1:dim(s_z)[1]){
  sd_sb[i] = sd(s_zb[i,])
}

par(mfrow=c(1,2))
plot(s_xb,s_zb,xlab=expression(beta),ylab=expression(S(z)))
abline(v=bcrit,col="red")

plot(beta,sd_sb,xlab=expression(beta),ylab=expression(sigma[S(z)]))
abline(v=bcrit,col="red")
```

```{r }
par(mfrow=c(1,2))
plot(s_x,s_z,xlab=expression(beta),ylab=expression(S(z)))
points(s_xr,s_zr,col = "blue")
points(s_xb,s_zb, col = "green")
abline(v=bcrit,col="red")

plot(beta,sd_s,xlab=expression(beta),ylab=expression(sigma[S(z)]))
points(beta,sd_sr,col = "blue")
points(beta,sd_sb, col = "green")
abline(v=bcrit,col="red")

```

the geometry seems not to affect the value of beta critic

# changing the order of the neighbourhood to second order

```{r }
mask <-matrix(0,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  mask[pixels[i,1],pixels[i,2]] = 1
}

neigh <- getNeighbors(mask = mask, c(2,2,2,2))
block <- getBlocks(mask = mask, 4)

```


```{r }
beta <- seq(0, 2, by = 0.05)
iter <- 1000
burn <- 200
n <- sum(mask)
k <- 3
bcrit <- log(1 + sqrt(k))
samp.bIS <- matrix(nrow = length(beta), ncol = iter - burn)
```


```{r }
for (i in 1:length(beta)) {
  result <- swNoData(beta[i],k,neigh,block,iter)
  samp.bIS[i,] <- result$sum[(burn+1):iter]
}
```


```{r }
s_z <- samp.bIS
s_x <- rep(beta,times=iter-burn)
s.frame <- data.frame(s_z,c(s_x,s_x))
names(s.frame) <- c("stat","beta")
sd_s <- rep(0,dim(s_z)[1])
for(i in 1:dim(s_z)[1]){
  sd_s[i] = sd(s_z[i,])
}

par(mfrow=c(1,2))
plot(s_x,s_z,xlab=expression(beta),ylab=expression(S(z)))
abline(v=bcrit,col="red")

plot(beta,sd_s,xlab=expression(beta),ylab=expression(sigma[S(z)]))
abline(v=bcrit,col="red")
```

**the value of beta critic is affected by he order of the beighbourhood, in his case the theoretical estimate is not valid **
