---
title: "Potts Model NO MASK"
author: Simone Colombara, Alessia Cotroneo, Francesco De Caro, Riccardo Morandi, Chiara
  Schembri, Alfredo Zapiola
date: "2022-11-05"
output: 
  html_document:
    toc: true
    toc_float: true
---
```{r libraries, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rayshader)
library(patchwork)
library(ggplot2)
library(plot.matrix)
library(bayesImageS)
library(coda)
library(viridis)
```

inporting data as usual

```{r}
D = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-IM-Step1-Step2-Step4-Step5-101.txt")
D0 = D
D0[is.na(D0)] = 0

pixels = read.table("/Users/macbookpro/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-XYCoordinates-Step1-Step2-Step4-Step5-101.txt")
colnames(D0) = substr(colnames(D0),1,4)
colnames(pixels) = c("x","y")

Data_long = as_tibble(data.frame( pixels, D0 ))
max_number_of_pixels = apply(Data_long[,1:2],2,max)

Data_very_long = reshape2::melt(Data_long,c("x","y")) %>% mutate(pixel_ind = paste0(x,"_",y), value_ind = rep(1:nrow(Data_long),ncol(D0)))
Data_very_long = Data_very_long %>% group_by(pixel_ind) %>% mutate(n = row_number()) %>% ungroup() %>% mutate(mz = as.numeric(substr(variable,2,4)))

```

# CONVENTOIONAL PCA ON VECTOR DATA 
**TO BE UPDATED ONCE WE FIX FPCA**

```{r}
pca = princomp(D0)
#auxiliary vector of pca components
pcascore1 <-as.vector(pca$scores[,1])

max(pca$score[,1])
min(pca$score[,1])

#sanity check
sum(is.na(pcascore1))
```

building auxiliary masks 

```{r}
mask <-matrix(0,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  mask[pixels[i,1],pixels[i,2]] = 1
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(mask, border=NA,asp = TRUE)

full_mask <-matrix(1,max_number_of_pixels[1],max_number_of_pixels[2])
```

building auxiliary pca vector with background class having pcascore1 = 100

```{r}
mask_idx = 1
pca_aux <-rep(100,max_number_of_pixels[1]*max_number_of_pixels[2])

for(i in 1:max_number_of_pixels[1]){
  for(j in 1:max_number_of_pixels[2]){
    if(mask[i,j]==1){
      pca_aux[(i-1)*max_number_of_pixels[2]+j] = pcascore1[mask_idx]
      mask_idx = mask_idx+1
    }
  }
}
```

sanity check

```{r}
pca_matr <- matrix(0,max_number_of_pixels[1],max_number_of_pixels[2])

for(i in 1:max_number_of_pixels[1]){
  for(j in 1:max_number_of_pixels[2]){
      pca_matr[i,j] = pca_aux[(i-1)*max_number_of_pixels[2]+j]
  }
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(pca_matr, border=NA,asp = TRUE,col = viridis,axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```

# K = 3

```{r}
q <- 4
betacritic = log(1 + sqrt(q))
neigh <- getNeighbors(mask = full_mask, c(2,2,0,0))
block <- getBlocks(mask = full_mask, 2)

priors <- list()
priors$k <- q
priors$mu <- c(-15,0,15,100)
priors$mu.sd <- c(rep(30,q-1),0.1)
priors$sigma <- c(rep(2,q-1),0.5)
priors$sigma.nu <- c(rep(0.125,q-1),1)
priors$beta <- c(0,betacritic)

mh <- list(algorithm="pseudolikelihood", bandwidth=1)

```


```{r, results='hide'}
res <- mcmcPotts(pca_aux, neigh, block, priors, mh, 10000, 5000)
```

## chain analysis 

```{r}
muchain = mcmc(res$mu)
varnames(muchain)<-c("mu_1","mu_2","mu_3","mu_b")
sigmachain = mcmc(res$sigma)
varnames(sigmachain)<-c("sigma_1","sigma_2","sigma_3","sigma_b")
betachain = mcmc(res$beta)
varnames(betachain)<-c("beta")
sumchain  = mcmc(res$sum)
varnames(sumchain)<-c("sum")
```

### mu

```{r}

summary(muchain)
batchSE(muchain)
effectiveSize(muchain)
rejectionRate(muchain)
par(mfrow=c(2,2))
plot(muchain,auto.layout=FALSE)
autocorr.plot(muchain,auto.layout=FALSE)



```

### sigma 

```{r}

summary(sigmachain)
batchSE(sigmachain)
effectiveSize(sigmachain)
par(mfrow=c(2,2))
plot(sigmachain,auto.layout=FALSE)
autocorr.plot(sigmachain,auto.layout=FALSE)
```

### sum

```{r}
summary(sumchain)
batchSE(sumchain)
effectiveSize(sumchain)

plot(sumchain)

autocorr.plot(sumchain)
```


### beta

```{r}
summary(betachain)

plot(betachain)

autocorr.plot(betachain)
```

this seems to want to go up but I fixed the threshold to beta critic to avoid the beta from exploding and giving all uniform parameters

## plot

```{r}
clustering <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:max_number_of_pixels[1]){
  for(j in 1:max_number_of_pixels[2]){
      clustering[i,j] = res$e[(i-1)*max_number_of_pixels[2]+j]
  }
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering, border=NA,asp = TRUE,col =viridis,axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```

```{r}
print(table(clustering))
```

**we have exactly 9717 pixel in the backround and they are all being identified as such** 

# removing the constraints on beta 
```{r, results='hide'}
priors2 <- list()
priors2$k <- q
priors2$mu <- c(-15,0,15,100)
priors2$mu.sd <- c(rep(30,q-1),0.01)
priors2$sigma <- c(rep(2,q-1),0.1)
priors2$sigma.nu <- c(rep(0.125,q-1),1)
priors2$beta <- c(0,5*betacritic)

mh2 <- list(algorithm="pseudolikelihood", bandwidth=1)

res2 <- mcmcPotts(pca_aux, neigh, block, priors2, mh2, 10000, 5000)

clustering2 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:max_number_of_pixels[1]){
  for(j in 1:max_number_of_pixels[2]){
      clustering2[i,j] = res2$e[(i-1)*max_number_of_pixels[2]+j]
  }
}

```

## chain analysis

```{r}

muchain2 = mcmc(res2$mu)
varnames(muchain2)<-c("mu_1","mu_2","mu_3","mu_b")
sigmachain2 = mcmc(res2$sigma)
varnames(sigmachain2)<-c("sigma_1","sigma_2","sigma_3","sigma_b")
betachain2 = mcmc(res2$beta)
varnames(betachain2)<-c("beta")
sumchain2  = mcmc(res2$sum)
varnames(sumchain2)<-c("sum")

```

### mu

```{r}

summary(muchain2)
batchSE(muchain2)
effectiveSize(muchain2)
rejectionRate(muchain2)

par(mfrow=c(2,2))

plot(muchain2,auto.layout=FALSE)

par(mfrow=c(2,2))
autocorr.plot(muchain2,auto.layout=FALSE)

```

### sigma

```{r}
summary(sigmachain2)
batchSE(sigmachain2)
effectiveSize(sigmachain2)

par(mfrow=c(2,2))
plot(sigmachain2,,auto.layout=FALSE)

par(mfrow=c(2,2))
autocorr.plot(sigmachain2,auto.layout=FALSE)
```

###sum

```{r}
summary(sumchain2)
batchSE(sumchain2)
effectiveSize(sumchain2)

plot(sumchain2)

autocorr.plot(sumchain2)
```

still needs thinning
batchSE broke for some reason
are the spikes in the traceplot indicating that something went wrong?

###beta

```{r}
summary(betachain2)
batchSE(betachain2)
effectiveSize(betachain2)

plot(betachain2)

autocorr.plot(betachain2)

```

this seems to want to go up but I fixed the threshold to beta critic to avoid the beta from exploding and giving all uniform parameters
the beta stabilizes at 2.5 which is above the beta critic

```{r}
print(table(clustering2))
```

## plot

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering2, border=NA,asp = TRUE,col =viridis,axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```


**the fact that beta is bigger than beta critic leads to this type of clusters**

**using the full matrix did not change anything**

# K = 6

```{r, results='hide'}
q <- 7
betacritic6 = log(1 + sqrt(q))
neigh <- getNeighbors(mask = full_mask, c(2,2,0,0))
block <- getBlocks(mask = full_mask, 2)

priors6 <- list()
priors6$k <- q
priors6$mu <- c(-23,-15,-7,0,7,15,100)
priors6$mu.sd <- c(rep(30,q-1),0.01)
priors6$sigma <- c(rep(2,q-1), 0.1)
priors6$sigma.nu <- c(rep(0.125,q-1),1)
priors6$beta <- c(0,betacritic6)

mh6 <- list(algorithm="pseudolikelihood", bandwidth=1)

res6 <- mcmcPotts(pca_aux, neigh, block, priors6, mh6, 10000, 5000)

clustering6 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:max_number_of_pixels[1]){
  for(j in 1:max_number_of_pixels[2]){
      clustering6[i,j] = res6$e[(i-1)*max_number_of_pixels[2]+j]
  }
}

```

## chain analysis

```{r}
muchain6 = mcmc(res6$mu)
varnames(muchain6)<-c("mu_1","mu_2","mu_3","mu_4","mu_5","mu_6","mu_b")
sigmachain6 = mcmc(res6$sigma)
varnames(sigmachain6)<-c("sigma_1","sigma_2","sigma_3","sigma_4","sigma_5","sigma_6","sigma_b")
betachain6 = mcmc(res6$beta)
varnames(betachain6)<-c("beta")
sumchain6  = mcmc(res6$sum)
varnames(sumchain6)<-c("sum")

```

### mu 

```{r}
summary(muchain6)
batchSE(muchain6)
effectiveSize(muchain6)
rejectionRate(muchain6)

par(mfrow=c(2,2))
plot(muchain6,auto.layout = FALSE)

par(mfrow=c(2,2))
autocorr.plot(muchain6,auto.layout=FALSE)

```

### sigma

```{r}
summary(sigmachain6)
batchSE(sigmachain6)
effectiveSize(sigmachain6)
par(mfrow=c(2,2))
plot(sigmachain6,auto.layout=FALSE)

autocorr.plot(sigmachain6,auto.layout=FALSE)
```

### sum

```{r}
summary(sumchain6)
batchSE(sumchain6)
effectiveSize(sumchain6)

plot(sumchain6)
autocorr.plot(sumchain6)
```

### beta

```{r}
summary(betachain6)
batchSE(betachain6)
effectiveSize(betachain6)

plot(betachain6)

autocorr.plot(betachain6)
```


## plot

```{r}
print(table(clustering6))
```

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering6, border=NA,asp = TRUE,col=viridis(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```


## removing the constraint on beta critic 

we needed to run more ierations in this case to have convergent chains

```{r, results='hide'}
priors6b <- priors6
priors6b$beta <- c(0,5*betacritic6)

res6b <- mcmcPotts(pca_aux, neigh, block, priors6b, mh6, 20000, 10000)

clustering6b <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:max_number_of_pixels[1]){
  for(j in 1:max_number_of_pixels[2]){
    clustering6b[i,j] = res6b$e[(i-1)*max_number_of_pixels[2]+j]
  }
}
```


```{r}
muchain6b = mcmc(res6b$mu)
varnames(muchain6b)<-c("mu_1","mu_2","mu_3","mu_4","mu_5","mu_6","mu_b")
sigmachain6b = mcmc(res6b$sigma)
varnames(sigmachain6b)<-c("sigma_1","sigma_2","sigma_3","sigma_4","sigma_5","sigma_6","sigma_b")
betachain6b = mcmc(res6b$beta)
varnames(betachain6b)<-c("beta")
sumchain6b  = mcmc(res6b$sum)
varnames(sumchain6b)<-c("sum")

```

### mu

```{r}
summary(muchain6b)
batchSE(muchain6b)

effectiveSize(muchain6b)
rejectionRate(muchain6b)
par(mfrow=c(2,2))
plot(muchain6b,auto.layout=FALSE)

autocorr.plot(muchain6b,auto.layout=FALSE)

```

### sigma

```{r}
summary(sigmachain6b)
batchSE(sigmachain6b)
effectiveSize(sigmachain6b)

par(mfrow=c(2,2))
plot(sigmachain6b,auto.layout=FALSE)

autocorr.plot(sigmachain6b,auto.layout=FALSE)
```

this needs to be thinned as well 
the chains look very strange and I don't know what to do about it

### sum
```{r}
summary(sumchain6b)
batchSE(sumchain6b)
effectiveSize(sumchain6b)

plot(sumchain6b)

autocorr.plot(sumchain6b)
```

other than high autocorrelation that can be fixed by thinning the chain

### beta

```{r}
summary(betachain6b)
batchSE(betachain6b)
effectiveSize(betachain6b)

plot(betachain6b)

autocorr.plot(betachain6b)
```

 we don't have the problem of rejecting like in the other chain
 beta also in this case goes to around 2 which is bigger than beta-critic6 = 1.238226
 
## plot
 
```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering6b, border=NA,asp = TRUE,col=turbo(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```
 this clearly suffers from the beta problem there is a dominationg class 
 
```{r}
 print(table(res6b$e))
```

**this is not working**

# refining the priors

since we have a lot of groups we try o reduce the prior variance in he hope of better distinguishing different groups

```{r, results='hide'}
q <- 7
betacritic6 = log(1 + sqrt(q))

priors6 <- list()
priors6$k <- q
priors6$mu <- c(-23,-15,-7,0,7,15,100)
priors6$mu.sd <- c(rep(10,q-1),0.01)
priors6$sigma <- c(rep(2,q-1),0.1)
priors6$sigma.nu <- c(rep(0.5,q-1),1)
priors6$beta <- c(0,5*betacritic6)

mh6 <- list(algorithm="pseudolikelihood", bandwidth=1)

res6 <- mcmcPotts(pca_aux, neigh, block, priors6, mh6, 15000, 10000)

clustering6c <-matrix(NA,max_number_of_pixels[2],max_number_of_pixels[2])
for(i in 1:max_number_of_pixels[2]){
  for(j in 1:max_number_of_pixels[2]){
      clustering6c[i,j] = res6$e[(i-1)*max_number_of_pixels[2]+j]
  }
}

```

## chain analysis

```{r}
muchain6 = mcmc(res6$mu)
varnames(muchain6)<-c("mu_1","mu_2","mu_3","mu_4","mu_5","mu_6","mu_b")
sigmachain6 = mcmc(res6$sigma)
varnames(sigmachain6)<-c("sigma_1","sigma_2","sigma_3","sigma_4","sigma_5","sigma_6","sigma_b")
betachain6 = mcmc(res6$beta)
varnames(betachain6)<-c("beta")
sumchain6  = mcmc(res6$sum)
varnames(sumchain6)<-c("sum")

```

### mu 

```{r}
summary(muchain6)
batchSE(muchain6)
effectiveSize(muchain6)
rejectionRate(muchain6)

par(mfrow=c(2,2))
plot(muchain6,auto.layout = FALSE)

par(mfrow=c(2,2))
autocorr.plot(muchain6,auto.layout=FALSE)

```

### sigma

```{r}
summary(sigmachain6)
batchSE(sigmachain6)
effectiveSize(sigmachain6)
par(mfrow=c(2,2))
plot(sigmachain6,auto.layout=FALSE)

autocorr.plot(sigmachain6,auto.layout=FALSE)
```

### sum

```{r}
summary(sumchain6)

plot(sumchain6)
autocorr.plot(sumchain6)
```

### beta

```{r}
summary(betachain6)

plot(betachain6)

autocorr.plot(betachain6)
```

## plot

```{r}
print(table(clustering6c))
```

```{r}
par(mar=c(4.1, 4.1, 4.1, 4.1))
plot(clustering6c, border=NA,asp = TRUE,col=turbo(q),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL)
```


# Final Plot

```{r}
par(mfrow=c(2,2))
plot(clustering, border=NA,asp = TRUE,col =viridis(4),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL,main = "K = 3 & betacritic")
plot(clustering2, border=NA,asp = TRUE,col =plasma(4),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL,main = "K = 3")
plot(clustering6, border=NA,asp = TRUE,col =magma(7),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL,main = "K = 6 & betacritic")
plot(clustering6c, border=NA,asp = TRUE,col=turbo(7),axis.col=NULL, axis.row=NULL, xlab='', ylab='',key = NULL,main = "K = 6")

```