---
title: "Potts_first_trial"
author: "Riccardo Morandi"
date: "2022-10-29"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r libraries, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rayshader)
library(patchwork)
library(ggplot2)
library(plot.matrix)
library(bayesImageS)
library(coda)
```

inporting data as usual


```{r}
D = read.table("/Users/macbook/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-IM-Step1-Step2-Step4-Step5-101.txt")
D0 = D
D0[is.na(D0)] = 0

pixels = read.table("/Users/macbook/Documents/Bayesian Statistics/Project/Raw_data/LIPIDI/78 variabili/101_lipidi-PreProcessed-XYCoordinates-Step1-Step2-Step4-Step5-101.txt")
colnames(D0) = substr(colnames(D0),1,4)
colnames(pixels) = c("x","y")

Data_long = as_tibble(data.frame( pixels, D0 ))
max_number_of_pixels = apply(Data_long[,1:2],2,max)

Data_very_long = reshape2::melt(Data_long,c("x","y")) %>% mutate(pixel_ind = paste0(x,"_",y), value_ind = rep(1:nrow(Data_long),ncol(D0)))
Data_very_long = Data_very_long %>% group_by(pixel_ind) %>% mutate(n = row_number()) %>% ungroup() %>% mutate(mz = as.numeric(substr(variable,2,4)))

```

# CONVENTOIONAL PCA ON VECTOR DATA 
**TO BE UPDATED ONCE WE FIX FPCA**

```{r}
pca = princomp(D0)
#auxiliary vector of pca components
pcascore1 <-as.vector(pca$scores[,1])

max(pca$score[,1])
min(pca$score[,1])

#sanity check
sum(is.na(pcascore1))
```

building auxiliary mask to pass junst the measured pixels to the functions

```{r}
mask <-matrix(0,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  mask[pixels[i,1],pixels[i,2]] = 1
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(mask, border=NA,asp = TRUE)
```

# K = 3

Specifying the number of clusters, the priors and the Metropolis-Hastings parameters

```{r}
q <- 3
betacritic = log(1 + sqrt(q))
neigh <- getNeighbors(mask = mask, c(2,2,0,0))
block <- getBlocks(mask = mask, 2)

priors <- list()
priors$k <- q
priors$mu <- c(-5,0,5)
priors$mu.sd <- rep(0.5,q)
priors$sigma <- rep(2,q)
priors$sigma.nu <- rep(1.5,q)
priors$beta <- c(0,betacritic)

mh <- list(algorithm="pseudolikelihood", bandwidth=1)
```

```{r, results='hide'}
res <- mcmcPotts(pcascore1, neigh, block, priors, mh, 10000, 5000)
```

## chain analysis 

```{r}
muchain = mcmc(res$mu)
varnames(muchain)<-c("mu_1","mu_2","mu_3")
sigmachain = mcmc(res$sigma)
varnames(sigmachain)<-c("sigma_1","sigma_2","sigma_3")
betachain = mcmc(res$beta)
varnames(betachain)<-c("beta")
sumchain  = mcmc(res$sum)
varnames(sumchain)<-c("sum")
```

### mu

```{r}

summary(muchain)
batchSE(muchain)
effectiveSize(muchain)
rejectionRate(muchain)
# still need to understand this since i am using adaptive mh alg
plot(muchain)

par(mfrow=c(3,1))
autocorr.plot(muchain,auto.layout=FALSE)

crosscorr.plot(muchain)
```

### sigma 

```{r}

summary(sigmachain)
batchSE(sigmachain)
effectiveSize(sigmachain)

plot(sigmachain)

par(mfrow=c(3,1))
autocorr.plot(sigmachain,auto.layout=FALSE)
```

### sum

```{r}
summary(sumchain)
batchSE(sumchain)
effectiveSize(sumchain)

plot(sumchain)

autocorr.plot(sumchain)
```

up to here the chain does not present mixing problems (this is probably due to the adaptiveness of the algorithm that prevents it) 
other than high autocorrelation that can be fixed by thinning the chain

### beta

```{r}
summary(betachain)
batchSE(betachain)
effectiveSize(betachain)

plot(betachain)

autocorr.plot(betachain)
```

this seems to want to go up but I fixed the threshold to beta critic to avoid the beta from exploding and giving all uniform parameters

## plot

```{r}
clustering <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering[pixels[i,1],pixels[i,2]] = res$e[i]
}

par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering, border=NA,asp = TRUE,col = rainbow)
```


## removing the constraints on beta 

```{r, results='hide'}
priors2 <- list()
priors2$k <- q
priors2$mu <- c(-5,0,5)
priors2$mu.sd <- rep(0.5,q)
priors2$sigma <- rep(2,q)
priors2$sigma.nu <- rep(1.5,q)
priors2$beta <- c(0,2*betacritic)

mh2 <- list(algorithm="pseudolikelihood", bandwidth=1)

res2 <- mcmcPotts(pcascore1, neigh, block, priors2, mh2, 10000, 5000)

clustering2 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering2[pixels[i,1],pixels[i,2]] = res2$e[i]
}
```

```{r}

muchain2 = mcmc(res2$mu)
varnames(muchain2)<-c("mu_1","mu_2","mu_3")
sigmachain2 = mcmc(res2$sigma)
varnames(sigmachain2)<-c("sigma_1","sigma_2","sigma_3")
betachain2 = mcmc(res2$beta)
varnames(betachain2)<-c("beta")
sumchain2  = mcmc(res2$sum)
varnames(sumchain2)<-c("sum")

```

### mu

```{r}

summary(muchain2)
batchSE(muchain2)
effectiveSize(muchain2)
rejectionRate(muchain2)
# still need to understand this since i am using adaptive mh alg

plot(muchain2)

par(mfrow=c(3,1))
autocorr.plot(muchain2,auto.layout=FALSE)

crosscorr.plot(muchain2)

```

### sigma

```{r}
summary(sigmachain2)
batchSE(sigmachain2)
effectiveSize(sigmachain2)

plot(sigmachain2)

par(mfrow=c(3,1))
autocorr.plot(sigmachain2,auto.layout=FALSE)
```

### sum

```{r}
summary(sumchain2)
batchSE(sumchain2)
effectiveSize(sumchain2)

plot(sumchain2)

autocorr.plot(sumchain2)
```

still needs thinning

### beta

```{r}
summary(betachain2)
batchSE(betachain2)
effectiveSize(betachain2)

plot(betachain2)

autocorr.plot(betachain2)

```

this seems to want to go up but I fixed the threshold to beta critic to avoid the beta from exploding and giving all uniform parameters
the beta stabilizes at 1.6 which is above the beta critic

## plot

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering, border=NA,asp = TRUE,col = rainbow)
```

I now will try with different values of q to see the difference in particular with q = 6 and 9 to compare with the kmeans baseline
I will try both the constrained  and not constrained beta

# K = 6

```{r, results='hide'}
q <- 6
betacritic6 = log(1 + sqrt(q))
neigh <- getNeighbors(mask = mask, c(2,2,0,0))
block <- getBlocks(mask = mask, 2)

priors6 <- list()
priors6$k <- q
priors6$mu <- c(-5,-2.5,0,2.5,5,7.5)
priors6$mu.sd <- rep(0.5,q)
priors6$sigma <- rep(2,q)
priors6$sigma.nu <- rep(1.5,q)
priors6$beta <- c(0,betacritic6)

mh6 <- list(algorithm="pseudolikelihood", bandwidth=1)

res6 <- mcmcPotts(pcascore1, neigh, block, priors6, mh6, 10000, 5000)

clustering6 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering6[pixels[i,1],pixels[i,2]] = res6$e[i]
}

```

## chain analysis

```{r}
muchain6 = mcmc(res6$mu)
varnames(muchain6)<-c("mu_1","mu_2","mu_3","mu_4","mu_5","mu_6")
sigmachain6 = mcmc(res6$sigma)
varnames(sigmachain6)<-c("sigma_1","sigma_2","sigma_3","sigma_4","sigma_5","sigma_6")
betachain6 = mcmc(res6$beta)
varnames(betachain6)<-c("beta")
sumchain6  = mcmc(res6$sum)
varnames(sumchain6)<-c("sum")

```

### mu 

```{r}
summary(muchain6)
batchSE(muchain6)
effectiveSize(muchain6)
rejectionRate(muchain6)


plot(muchain6)

par(mfrow=c(2,3))
autocorr.plot(muchain6,auto.layout=FALSE)

crosscorr.plot(muchain6)
```

### sigma

```{r}
summary(sigmachain6)
batchSE(sigmachain6)
effectiveSize(sigmachain6)

plot(sigmachain6)

par(mfrow=c(2,3))
autocorr.plot(sigmachain6,auto.layout=FALSE)
```

### sum

```{r}
summary(sumchain6)
batchSE(sumchain6)
effectiveSize(sumchain6)

plot(sumchain6)
autocorr.plot(sumchain6)
```

### beta

```{r}
summary(betachain6)
batchSE(betachain6)
effectiveSize(betachain6)

plot(betachain6)

autocorr.plot(betachain6)
# there are a lot of mixing problems with the chain, i am rejecting all the time in mh
#res6$accept is 54 on 10000 samples keeping in mind that we used an adaptive alg and the final bandwidth is 0.03603595
```

there are a lot of mixing problems with the chain, i am rejecting all the time in mh
res6$accept is 54 on 10000 samples keeping in mind that we used an adaptive alg and the final bandwidth is 0.03603595

## plot

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering6, border=NA,asp = TRUE,col = rainbow)
```


## removing the constraint on beta critic 

```{r, results='hide'}
priors6b <- priors6
priors6b$beta <- c(0,2*betacritic6)

res6b <- mcmcPotts(pcascore1, neigh, block, priors6b, mh6, 10000, 5000)

clustering6b <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering6b[pixels[i,1],pixels[i,2]] = res6b$e[i]
}
```


```{r}
muchain6b = mcmc(res6b$mu)
varnames(muchain6b)<-c("mu_1","mu_2","mu_3","mu_4","mu_5","mu_6")
sigmachain6b = mcmc(res6b$sigma)
varnames(sigmachain6b)<-c("sigma_1","sigma_2","sigma_3","sigma_4","sigma_5","sigma_6")
betachain6b = mcmc(res6b$beta)
varnames(betachain6b)<-c("beta")
sumchain6b  = mcmc(res6b$sum)
varnames(sumchain6b)<-c("sum")

```

### mu

```{r}
summary(muchain6b)
batchSE(muchain6b)

effectiveSize(muchain6b)
rejectionRate(muchain6b)

plot(muchain6b)

par(mfrow=c(2,3))
autocorr.plot(muchain6b,auto.layout=FALSE)

crosscorr.plot(muchain6b)
# there are weird jumps int he chain i don't know why
```

### sigma

```{r}
summary(sigmachain6b)
batchSE(sigmachain6b)
effectiveSize(sigmachain6b)

plot(sigmachain6b)

par(mfrow=c(2,3))
autocorr.plot(sigmachain6b,auto.layout=FALSE)
```

this needs to be thinned as well 

### sum

```{r}
summary(sumchain6b)
batchSE(sumchain6b)
effectiveSize(sumchain6b)

plot(sumchain6b)

autocorr.plot(sumchain6b)
```

other than high autocorrelation that can be fixed by thinning the chain

### beta

```{r}
summary(betachain6b)
batchSE(betachain6b)
effectiveSize(betachain6b)

plot(betachain6b)

autocorr.plot(betachain6b)
```

 we don't have the problem of rejecting like in the othe chain
 beta also in this case goes to arround 2 which is bigger than betacritic6 = 1.238226
 
## plot
 
```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering6b, border=NA,asp = TRUE,col = rainbow)
```
 this clearly suffers from the beta problem there is a dominationg class 
 
```{r}
 print(table(res6b$e))
```

# K = 9

```{r, results='hide'}
q <- 9
betacritic9 = log(1 + sqrt(q))

priors9 <- list()
priors9$k <- q
priors9$mu <- c(-8,-6,-4,-2,0,2,4,6,8)
priors9$mu.sd <- rep(0.5,q)
priors9$sigma <- rep(2,q)
priors9$sigma.nu <- rep(1.5,q)
priors9$beta <- c(0,betacritic9)

mh9 <- list(algorithm="pseudolikelihood", bandwidth=1)

res9 <- mcmcPotts(pcascore1, neigh, block, priors9, mh9, 10000, 5000)

clustering9 <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering9[pixels[i,1],pixels[i,2]] = res9$e[i]
}
```

## chain analysis

```{r}
muchain9 = mcmc(res9$mu)
varnames(muchain9)<-c("mu_1","mu_2","mu_3","mu_4","mu_5","mu_6","mu_7","mu_8","mu_9")
sigmachain9 = mcmc(res9$sigma)
varnames(sigmachain9)<-c("sigma_1","sigma_2","sigma_3","sigma_4","sigma_5","sigma_6","sigma_7","sigma_8","sigma_9")
betachain9 = mcmc(res9$beta)
varnames(betachain9)<-c("beta")
sumchain9  = mcmc(res9$sum)
varnames(sumchain9)<-c("sum")

```

###mu 

```{r}
summary(muchain9)
batchSE(muchain9)
effectiveSize(muchain9)
rejectionRate(muchain9)

plot(muchain9)
par(mfrow=c(3,3))
autocorr.plot(muchain9,auto.layout=FALSE)

crosscorr.plot(muchain9)

```


### sigma

```{r}
summary(sigmachain9)
batchSE(sigmachain9)
effectiveSize(sigmachain9)

plot(sigmachain9)

par(mfrow=c(3,3))
autocorr.plot(sigmachain9,auto.layout=FALSE)
```

### sum

```{r}
summary(sumchain9)
batchSE(sumchain9)
effectiveSize(sumchain9)

plot(sumchain9)

autocorr.plot(sumchain9)
```

### beta

```{r}
summary(betachain9)
batchSE(betachain9)
effectiveSize(betachain9)

plot(betachain9)

autocorr.plot(betachain9)
```

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering9, border=NA,asp = TRUE,col = topo.colors(q))
```

the clustering kind of works

## plot

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering9, border=NA,asp = TRUE, col = topo.colors(q))
# same problem as before since the beta is above beta critic
```

```{r}
print(table(res9$e))
```

## removing the constriant on beta

```{r,  results='hide'}
priors9b <- priors9
priors9b$beta <- c(0,2*betacritic9)

res9b <- mcmcPotts(pcascore1, neigh, block, priors9b, mh9, 10000, 5000)

clustering9b <-matrix(NA,max_number_of_pixels[1],max_number_of_pixels[2])
for(i in 1:dim(pixels)[1]){
  clustering9b[pixels[i,1],pixels[i,2]] = res9b$e[i]
}
```

## chain analysis

```{r}
muchain9b = mcmc(res9b$mu)
varnames(muchain9b)<-c("mu_1","mu_2","mu_3","mu_4","mu_5","mu_6","mu_7","mu_8","mu_9")
sigmachain9b = mcmc(res9b$sigma)
varnames(sigmachain9b)<-c("sigma_1","sigma_2","sigma_3","sigma_4","sigma_5","sigma_6","sigma_7","sigma_8","sigma_9")
betachain9b = mcmc(res9b$beta)
varnames(betachain9b)<-c("beta")
sumchain9b  = mcmc(res9b$sum)
varnames(sumchain9b)<-c("sum")
```

### mu
```{r}
summary(muchain9b)
batchSE(muchain9b)

effectiveSize(muchain9b)
rejectionRate(muchain9b)
par(mfrow=c(3,3))
plot(muchain9b)

autocorr.plot(muchain9b,auto.layout=FALSE)

crosscorr.plot(muchain9b)
# there are weird jumps int he chain i don't know why

```

### sigma

```{r}
summary(sigmachain9b)
batchSE(sigmachain9b)
effectiveSize(sigmachain9b)
plot(sigmachain9b)

par(mfrow=c(3,3))
autocorr.plot(sigmachain9b,auto.layout=FALSE)
```


### sum
```{r}
summary(sumchain9b)
batchSE(sumchain9b)
effectiveSize(sumchain9b)

plot(sumchain9b)

autocorr.plot(sumchain9b)
```

this chain has a lot of problems

### beta
```{r}
summary(betachain9b)
batchSE(betachain9b)
effectiveSize(betachain9b)

plot(betachain9b)

autocorr.plot(betachain9b)
```
we don't have the problem of rejecting like in the other chain
beta also in this case goes to arround 

## plot

```{r}
par(mar=c(5.1, 4.1, 4.1, 4.1))
plot(clustering9b, border=NA,asp = TRUE, col = topo.colors(q))
# same problem as before since the beta is above beta critic
```

same problem as before since the beta is above beta critic

```{r}
print(table(clustering9b))
```

